### 2021-10-05 _bookdown_files/04-results.Rmd
Working on the discussion

---
bibliography: references.bib
---

```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

```{r packages_data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

#Load required packages :
library(tidyverse)
library(readr)
library(scales)
library(ggplot2)
library(ggpubr)
library(sunburstR)
library(d3r)

#Load the data from the .csv files
source("./scripts/read_data.R", local = knitr::knit_global())

```

# Results

This section summarizes the findings relevant to our aforementioned research questions. \@ref(tab:RefTabQuant), \@ref(tab:RefTabQual) and \@ref(tab:RefTabRev) in our supplementary material, summarize every reference identified and the variables extracted to answer our research questions for quantitative studies, qualitative studies and reviews, respectively.

## Literature search

The electronic search identified 769 references and our handsearch of reference lists resulted in the addition of the seminal study by Stall and colleague @Stall2003 . After removing duplicates, 277 records were screened for inclusion.

After screening the title and abstract, 109 records were excluded. The two main reasons for exclusion at this stage were that MSM was not the main study population (n=46) and that syndemic was not the main focus of the paper (n=34). 29 studies were excluded because the type of publication did not meet our inclusion criteria.

The full texts of the 168 remaining references were obtained and read; 59 were excluded after this phase. During the previous step, we had decided to tentatively include papers in which the sample was not entirely comprised of MSM. After reviewing the full paper, we decided to exclude studies if the MSM data was aggregated with data from another population (n=9), most commonly transgender women. The rationale for this decision was to keep the focus clearly on MSM.

The PRISMA flowchart generated by DistillerSR can be found in Figure \@ref(fig:PRISMA)

```{r PRISMA, fig.cap= "PRISMA flowchart of the search strategy", fig.height=7.5, fig.id="PRISMA", fig.width=7, fig.align='center'}
knitr::include_graphics("Files/images/PRISMA.png")

```

## How are studies concerning Syndemic Theory applied to MSM conducted?

### General characteristics of included studies

```{r general_characteristics_data_preparation, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

#Source the scripts used to manipulate the data from the .csv files
source("./scripts/create_year_design.R", local = knitr::knit_global())
source("./scripts/create_location_data.R", local = knitr::knit_global())
source("./scripts/create_sample.R", local = knitr::knit_global())

#Create the plot 
source("./scripts/create_plot_year.R", local = knitr::knit_global())

```

The publication date of the studies ranges from `r min(year_design$year)` to `r max(year_design$year)` with a marked increase in the number papers published annually from 2013 onward as illustrated in Figure \@ref(fig:PlotYear). There is also an increase in the diversity of the studies, both in the study design used and the continents where the studies were conducted.

```{r plot_year, fig.cap="Plot of the cumulative number of studies published yearly", fig.height=7, fig.id="PlotYear", fig.width=5.5, fig.align='center'}

knitr::include_graphics("./Output/Plots/Plot_Year.png")

```

However, despite the increasing diversity, the vast majority of the studies were conducted in North America (N = `r .North_America_studies`), with a large predominance of studies in the United States of America (N= `r .USA_studies`). The majority of the remaining studies were then located in Asia (N = `r .Asia_studies`) or in Europe (N = `r .Europe_studies`). South America (N = `r .South_America_studies`) and Africa (N = `r .Africa_studies`)were the least represented continents.

In terms of design, a similar pattern appears, with most of the studies employing a cross-sectional design (N = `r .N_cross_sec`). The number of longitudinal studies has grown steadily since 2014 and represents `r percent(.N_long/nrow(overview_of_studies))` of the papers. The duration of longitudinal studies ranged from `r min(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` to `r max(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` months.

Compared to quantitative studies, qualitative studies are under-represented, with only `r .N_qual` papers. In terms of analysis method, the most common was thematic (content) analysis (N= `r sum(!is.na(summary_qualitative$thematic_content)) + sum(!is.na(summary_qualitative$thematic))`). The rest of the studies consisted in framework analysis (N = `r sum(!is.na(summary_qualitative$framework))`), grounded theory (N = `r sum(!is.na(summary_qualitative$grounded))`), analytic induction (N = `r sum(!is.na(summary_qualitative$analytic_induction))`) and constant comparative analysis (N = `r sum(!is.na(summary_qualitative$comparative))`).

We only identified 2 papers employing mixed methods design [@buttram2015; @Halkitis2012]; Buttram et al. used a cross-sectional design for the quantitative part and in-depth interviews analyzed through grounded theory for the qualitative part while Halkitis et al used a cross-sectional quantitative survey and in-depth discovery interviews.

The first systematic reviews and meta-analyses were published, in `r min(year_design$year[year_design$design == "Systematic reviews, meta-analyses"])`, while the first interventional studies were published in `r min(year_design$year[year_design$design == "Interventional"])`.

Finally, the median sample size of the studies was `r median(sample$sample_size)` (range : `r min(sample$sample_size)` - `r number(max(sample$sample_size))`) and the mean age of the participants, when reported, ranged from `r min(sample$age, na.rm = T)` to `r round(max(sample$age, na.rm = T), digits = 0)` years.

### Subpopulations studied

```{r Subpopulation_data_prep, message=FALSE, warning=FALSE, include=FALSE}
source("./scripts/create_population_data.R", local = knitr::knit_global())
source("./scripts/create_population_plot.R", local = knitr::knit_global())

```

Fewer than half of the studies (N = `r sum(!is.na(population$subpopulation))` ) focused on a subpopulation of MSM, of those, we identified 7 types of MSM subpopulations : (a) young MSM ; (b) older MSM ; (c) MSM from a racial/ethnic minority ; (d) MSM living with HIV ; (e) MSM engaged in sex work ; (f) Men who have Sex with Men and Women (MSMW) ; and (g) transgender MSM.

Among studies focusing on a specific age group, `r sum(population$age == "Young adult")` studies focus on young adults, `r sum(population$age == "Older men")` studies focus on older MSM[@zepf2020; @Halkitis2012] and `r sum(population$age == "Adolescent")` study focuses on adolescents[@Perry]. Of note, the two studies on older MSM are both entirely composed of a sample of older MSM living with HIV.

Black MSM (BMSM) are the most studied MSM from a racial/ethnic minority (N= `r sum(population$race_ethnicity == "Black")`) but studies focusing on them still represent only `r percent(sum(population$race_ethnicity == "Black")/104)` of our sample. Among these studies, `r .N_YBMSM` studies are composed of young BMSM[@Maiorana; @Reed2016] and `r .N_MSM_SW` study focuses on Black MSM engaged in sex work[@chandler2020].\
`r sum(population$race_ethnicity == "Latino")` studies focus on Latino MSM (LMSM) and `r .N_LMSMW` of these studies focus on Latino MSMW[@Munoz-Laboy2018]. Furthermore, `r sum(population$race_ethnicity == "Latino & Black")` study focus on both BMSM and LMSM[@cassels2020].

Concerning the serologic status of MSM in syndemic literature, `r sum(population$serologic_status == "HIV-positive")` studies are entirely comprised of a sample of MSM living with HIV. Apart from the studies mentioned above on older MSM living with HIV, `r sum(population$serologic_status == "HIV-positive" & population$age == "Young adult")` studies focus on young MSM living with HIV[@Lyons2013; @Bruce2011] . Notably, we didn't identify any studies focusing on racial minority MSM living with HIV.

Finally, the least represented subpopulation of MSM in this review are MSMW (N = `r sum(population$sex_behav == "MSMW")`), MSM engaged in sex work (N = `r sum(population$sex_work == "Sex workers")`) and transgender MSM (N = `r sum(population$gender_identity == "Transgender")`). Strikingly, transgender MSM, in addition to being the focus of only one study [@Reisner2016a] are, by design, excluded from every other study whereas other MSM subpopulation are at least partially represented in most of the samples.

Figure \@ref(fig:PlotPop) gives a visual representation of MSM subpopulation's representation in the studies included in this paper.

In order to obtain the most comprehensive knowledge of the representation of MSM subpopulations in syndemic literature, we also took into account studies that did not focus on any subpopulation in particular but who presented disaggregated data and analyses for one or more MSM subpopulations.\
`r .N_disaggregated` studies presented such data : `r .N_dis_BMSMW` for Black MSMW[@dyer2020], `r .N_dis_MSMW` for MSMW[@branstrom; @Mustanski2014; @friedman2014; @ferlatte2018], `r .N_dis_HIVMSM` for MSM living with HIV[@Kurtz2012; @ng2020] and `r .N_dis_LBMSM` for Latino and Black MSM[@Mustanski2017].

```{r plot_population, fig.cap="Number of studies focusing on a MSM subpopulation", fig.height=7, fig.id="PlotPop", fig.width=5.5, fig.align='center'}

knitr::include_graphics("./Output/Plots/Plot_population.png")
```

### Syndemic conditions and their measurement

#### Number and type of syndemic conditions identified

```{r syndemic_network_script, include=FALSE}

source("./scripts/create_syndemic_network.R", local = knitr::knit_global())

```

We identified `r nrow(nodes)` different syndemic conditions in the `r nrow(summary_quantitative)` quantitative studies we included : depression or depressive symptoms (N = `r nodes$studies[nodes$label == "depression"]`), intimate partner violence [IPV] (N = `r nodes$studies[nodes$label == "IPV"]`), substance use (N = `r nodes$studies[nodes$label == "substance_use"]`), childhood sexual abuse [CSA] (N = `r nodes$studies[nodes$label == "CSA"]`), polysubstance use (N = `r nodes$studies[nodes$label == "polysubstance"]`) , binge drinking (N = `r nodes$studies[nodes$label == "binge"]`), sexual compulsivity (N = `r nodes$studies[nodes$label == "sex_comp"]`), alcohol use disorder [AUD] (N = `r nodes$studies[nodes$label == "AUD"]`), suicidal thoughts and/or attempts (N = `r nodes$studies[nodes$label == "suicide"]`), sexual risk behaviors (N = `r nodes$studies[nodes$label == "sexual_risk"]`), experiences of violence (N = `r nodes$studies[nodes$label == "violence"]`), anxiety (N = `r nodes$studies[nodes$label == "anxiety"]`), substance use disorder [SUD] (N = `r nodes$studies[nodes$label == "substance_use_disorder"]`), alcohol use (N = `r nodes$studies[nodes$label == "alcohol_use"]`), discrimination (N = `r nodes$studies[nodes$label == "discrimination"]`), post-traumatic stress disorder [PTSD] (N = `r nodes$studies[nodes$label == "CSA"]`), chemsex (N = `r nodes$studies[nodes$label == "chemsex"]`), loneliness (N = `r nodes$studies[nodes$label == "loneliness"]`), incarceration (N = `r nodes$studies[nodes$label == "incarceration"]`), unstable housing (N = `r nodes$studies[nodes$label == "unstable_housing"]`), tobacco use (N = `r nodes$studies[nodes$label == "tobacco_use"]`), general mental distress (N = `r nodes$studies[nodes$label == "distress"]`), low social support (N = `r nodes$studies[nodes$label == "social_support"]`), low self-esteem (N = `r nodes$studies[nodes$label == "esteem"]`), internalised homophobia (N = `r nodes$studies[nodes$label == "internalised_homophobia"]`), exchange sex (N = `r nodes$studies[nodes$label == "exchange_sex"]`), sexually transmitted infections [STI] (N = `r nodes$studies[nodes$label == "STI"]`), childhood abuse (N = `r nodes$studies[nodes$label == "childhood"]`), intravenous drug use [IDU] (N = `r nodes$studies[nodes$label == "IDU"]`), poverty (N = `r nodes$studies[nodes$label == "poverty"]`), unemployment (N = `r nodes$studies[nodes$label == "unemployment"]`), poor healthcare access (N = `r nodes$studies[nodes$label == "healthcare_access"]`), involuntary subordination (N = `r nodes$studies[nodes$label == "subordination"]`), HIV diagnosis (N = `r nodes$studies[nodes$label == "HIV"]`), hostility (N = `r nodes$studies[nodes$label == "hostility"]`), stress (N = `r nodes$studies[nodes$label == "stress"]`), experience of trauma (N = `r nodes$studies[nodes$label == "trauma"]`), alexithymia (N = `r nodes$studies[nodes$label == "alexithymia"]`), attention deficit hyperactivity disorder [ADHD] (N = `r nodes$studies[nodes$label == "ADHD"]`), cognitive escape (N = `r nodes$studies[nodes$label == "escape"]`), frequenting gay social venues (N = `r nodes$studies[nodes$label == "venues"]`), hypersexuality (N = `r nodes$studies[nodes$label == "hypersexuality"]`), impulsivity (N = `r nodes$studies[nodes$label == "impulsivity"]`), poor physical health (N = `r nodes$studies[nodes$label == "physical_health"]`), school bullying (N = `r nodes$studies[nodes$label == "bullying"]`), and sleep disturbance (N = `r nodes$studies[nodes$label == "sleep_disturbance"]`).

We chose to distinguish between alcohol use disorder, binge drinking and alcohol use instead of merging them into a broad "alcohol-related syndemic condition". Some authors also considered "heavy drinking" [@Martinez2016a; @martinez2020] or "heavy alcohol use"[@Mimiaga2015b] but we chose to aggregate these conditions under "binge drinking" for clarity. Similarly, we distinguish between substance use, substance use disorder, intravenous drug use, polysubstance use, marijuana use, tobacco use and chemsex since they differ in terms of potential harm and the context of use.

```{r Stall_conditions, include=FALSE}
.Stall_conditions <- summary_quantitative %>%
  filter(!is.na(CSA) | (!is.na(depression) | (!is.na(polysubstance) | (!is.na(IPV)))))

```

Despite the high number of syndemic conditions studied in the literature, the impact of the seminal study by Stall et al.[@Stall2003] is striking. The conditions studied in that paper (depression, IPV, CSA and polysubstance use) are among the five most studied conditions in the field. Furthermore, `r percent(nrow(.Stall_conditions)/nrow(summary_quantitative))` of the quantitative studies included in our review consider at least one of the four conditions studied by Stall and colleagues as part of a syndemic.

#### Network analysis of the syndemic conditions

In order to better understand the connections between the syndemic conditions in the studies we elaborated a network plot (Figure \@ref(fig:SyndemicNetwork)). Nodes represent the syndemic conditions studied in the literature and edges (connectors) are drawn between two nodes when two conditions are studied in the same research paper. Furthermore, edges between nodes are thicker if the two conditions are frequently studied together. Furthermore, the outline of the nodes represent pie chart of their frequency in syndemic literature. For example, depression, the most studied condition, appears in `r percent((nodes$studies[nodes$label == "depression"])/nrow(summary_quantitative))` of our sample. As such, the node is circled by a chart representing this proportion. Moreover, as a force-directed graph, the location of the nodes gives an indication regarding their importance in terms of frequency and linkage to other conditions. As such, depression occupies a more central position than ADHD which was only studied once.

In this figure, we divided the conditions studied into 4 categories : mental health (e.g. depression, anxiety, PTSD) in blue, social conditions (e.g. substance use, loneliness) in teal, structural conditions (e.g. unemployment, healthcare access) in orange and physical health (e.g. STI, HIV) in red . The figure shows that structural syndemic conditions and physical health-related syndemic conditions are much less central, therefore much less studied, than social syndemic conditions and mental health-related syndemic conditions which are located towards the center of the figure.

Regarding the edges of the network, we identified `r nrow(edges)` pairs of syndemic conditions. However, nearly half of these pairs of conditions (N = `r nrow(.edges_1)`) appear in only one research paper. On the other hand , the most studied pairs of syndemic conditions were depression and IPV (N = `r edges$weight[edges$from == 1 & edges$to == 4]`), depression and substance use (N = (N = `r edges$weight[edges$from == 1 & edges$to == 2]`), depression and CSA (N = `r edges$weight[edges$from == 1 & edges$to == 10]`), depression and polysubstance use (N = `r edges$weight[edges$from == 1 & edges$to == 11]`) and IPV and CSA (N = `r edges$weight[edges$from == 4 & edges$to == 10]`).

In order to obtain a deeper understanding of the relation between the syndemic conditions, we also computed centrality indices for this network, namely degree centrality, the sum of weight in the network, closeness centrality, the inverse of the total length of the paths from a node to all other nodes, and betweenness, the number of shortest path passing through a node[@opsahl2010].\
Because the weight of the nodes in this network corresponds to the number of studies in which the syndemic conditions appear, degree centrality of each node is nearly perfectly correlated to the number of studies in which the syndemic condition was studied (`r paste0("r(", .cor_degree$parameter,")", " = ", round(.cor_degree$estimate, digits = 2), ", ", pvalue(.cor_degree$p.value, accuracy = 0.01, decimal.mark = ".", add_p = TRUE))`).\
Closeness centrality and betweenness centrality are also strongly correlated to the number of studies `r paste0("(r(", .cor_closeness$parameter,")", " = ", round(.cor_closeness$estimate, digits = 2), ", ", pvalue(.cor_closeness$p.value, accuracy = 0.01, decimal.mark = ".", add_p = TRUE), " and r(", .cor_betweenness$parameter,")", " = ", round(.cor_betweenness$estimate, digits = 2), ", ", pvalue(.cor_betweenness$p.value, accuracy = 0.01, decimal.mark = ".", add_p = TRUE), " respectively)")`. Interestingly, the betweenness centrality of unstable housing is notably high given the low number of studies focusing on this syndemic condition (N = `r nodes$studies[nodes$label == "unstable_housing"]`). Betweenness centrality can be seen as the extent of brokerage a node can exert on a network [@zweig2016]. Moreover in \@ref(fig:SyndemicNetwork) we see that the three structural syndemic conditions "unstable housing", "incarceration" and "poverty" are grouped together, indicating that they are frequently studied together. In other words, in syndemic literature, unstable housing seems to act as a bridge between the two aforementioned structural syndemic conditions and the rest of the syndemic conditions. Indeed, unstable housing has been studied alongside `r nrow(.edges_UH)` other syndemic conditions : depression, substance use, IPV, binge drinking, sexual compulsivity, IDU, CSA, polysubstance use, incarceration, poverty, unemployment, discrimination, poor healthcare access, alcohol use and violence.

```{r syndemic_network_plot, fig.cap="Network of the syndemic conditions studied in quantitative research. As a force-directed network, the most studied and interconnected conditions are the most central. Moreover, each node is circled by a pie chart of their frequency in the review. The thickness of the edges linking two nodes reflects the number of studies in which the two conditions were studied together.", fig.id="SyndemicNetwork", fig.width=7.45, fig.height=5.32, fig.align='center'}

knitr::include_graphics("./Output/Plots/syndemic_network.png")
```

#### Measurement of the syndemic conditions

```{r measurement_data, message=FALSE, warning=FALSE, include=FALSE}

source("./scripts/supplementary_tables/depression.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/IPV.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/substance.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/CSA.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/polysubstance.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/binge.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/violence.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/anxiety.R", local = knitr::knit_global())

```

The next part of our research question was to examine the way the syndemic conditions were studied in the literature. For the most often studied syndemic conditions, we found a high degree of variability in their measurement. Different authors used different scales or criteria and, even when the same scale or criteria was used, the cutoff or reference period differed among studies. For the sake of brevity, we will not provide a detailed description of the measurement method of the syndemic conditions studied in fewer than 10% of our sample. Detailed information about each of the papers included is presented in the supplementary material, part B ( \@ref(tab:DepressionTab) to \@ref(tab:AnxietyTab)).

For depression, `r length(unique(.depression_measurement$depression_scale_criteria))` different scales or criteria were used among the `r nodes$studies[nodes$label == "depression"]` studies in which this condition was studied.\
66 studies used scales while 10 studies used criteria to assess the presence of depression or depressive symptoms among the participants. The most frequently used scales were the full Center for Epidemiologic Studies-Depression scale [CESD] (N = 24) and the 10-items version of this scale [CESD-10] (N=10), the 9-items version of the Patient Health Questionnaire [PHQ-9] (N=8) and the Beck Depression Inventory [BDI-II] (N = 5).\
Cutoff for these scales varied the most for the CESD (range : 16 to 27) and the PHQ-9 (range : 5 to 15).\
When a scale was not used, the most frequent criteria to assess the presence of depression was the participant being on medication for depression in the past 12 months (N = 4).

For IPV, 5 types of intimate partner violence were identified : physical violence (N = `r .IPV_prevalence$sum[["physical IPV"]]`), sexual violence (N = `r .IPV_prevalence$sum[["sexual IPV"]]`), psychological violence (N = `r .IPV_prevalence$sum[["psychological IPV"]]`), gay-related violence (e.g. threats to reveal the partner's sexual orientation ; N = `r .IPV_prevalence$sum[["gay-related IPV"]]`) and HIV-related violence (e.g. threats to disclose the partner's serologic status ; N = `r .IPV_prevalence$sum[["HIV-related IPV"]]`). Reference periods varied from past month to lifetime.\
Furthermore, 7 studies used a scale to assess the presence of IPV : the HITS scale (N = 3) the Revised Conflict Tactics Scale [CTS2] (N= 3) and a scale developed by the authors of the study, assessing physical and psychological violence[@Yu2013].

For substance use, we looked at the class of substances explicitly screened by the authors and the reference period used. The most frequently screened class of substance was stimulants (i.e. cocaine/crack, amphetamines ; N = `r .substances_prev$sum[["stimulants"]]`), followed by depressants (i.e. GHB/GBL, benzodiazepines without prescriptions ; (N = `r .substances_prev$sum[["depressants"]]`), opioids (i.e. opioids misuse and ecstasy ; N = `r .substances_prev$sum[["opioids"]]`), marijuana (N = `r .substances_prev$sum[["marijuana"]]`), ecstasy /MDMA (N = `r .substances_prev$sum[["ecstasy"]]`), hallucinogens (i.e. ketamine, psilocybine, phencyclidine ; N = `r .substances_prev$sum[["hallucinogens"]]`), inhalants (nitrous oxyde, Popper ; N = `r .substances_prev$sum[["inhalants"]]`) and new psychoactive substances (i.e. synthetic cannabinoids, cathinones ; N = `r .substances_prev$sum[["NPS"]]`).\
Reference period ranged from the past month to over the lifetime.

For CSA, we identified `r length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Definition"]))` different definitions and `r length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Scale"]))` scales among the `r nodes$studies[nodes$label == "CSA"]` studies. The various definitions differed on the age at which the sexual intercourse occurred to be qualified as childhood sexual abuse, the age gap between the victim and the perpetrator and whether the sexual intercourse was unwanted. The most frequent definition was Finkelhor's definition[@finkelhor1994] : any sexual experience before 13 years old with someone at least 5 years older or any sexual experience between 13 and 17 years old with someone at least 10 years older (N = 7).

For polysubstance use, 17 studies considered that there was polysubstance use when 3 or more classes of substances were used while 9 studies defined polysubstance use as more than one class of substance being used. Furthermore, 4 studies excluded marijuana from the class of substances screened and 1 study excluded stimulants, because stimulants were already screened in another syndemic condition.

For binge drinking, we identified `r length(unique(.binge_measurement$binge_episodes))` different thresholds for binge drinking, differing on the number and frequency of binge drinking episodes. The threshold ranged from one episode in the past 12 months to more than one episode per week, every week in the past 12 months.

Measurement of sexual compulsivity was relatively standardized, compared to the rest of the syndemic conditions. Nearly every study used the Kalichman's Sexual Compulsivity Scale, with cutoff ranging from 20 to 26, except for two studies using the Compulsive Sexual Behavior Inventory[@Herrick2013; @Dyer2012] and one study using a scale devised by the authors @Wang2017.

For AUD, we identified 4 different definitions. `r sum(str_count(measurement$AUD_scale, pattern = "AUDIT-10"), na.rm = T)` studies used the full Alcohol Use Disorder Identification Test (AUDIT-10), `r sum(str_count(measurement$AUD_scale, pattern = "AUDIT-C"), na.rm = T)` studies used the 3-items version of this screening test (AUDIT-C), `r sum(str_count(measurement$AUD_scale, pattern = "CAGE"), na.rm = T)` studies used the CAGE questionnaire and one study used clinical diagnosis based on the DSM-IV criteria.

Regarding suicide, `r sum(is.na(measurement$suicide_attempts) & grepl(pattern = "suicidal_thoughts", measurement$suicidal_thoughts))` studies considered only suicidal thoughts, `r sum(grepl(pattern = "suicide_attempts", measurement$suicide_attempts) & grepl(pattern = "suicidal_thoughts", measurement$suicidal_thoughts))` studies considered both suicidal thoughts and suicide attempts and `r sum(grepl(pattern = "suicide_attempts", measurement$suicide_attempts) & is.na(measurement$suicidal_thoughts))` studies considered only suicide attempts. Furthermore, one study used the Positive and Negative Suicide Ideation scale (PANSI) and one study used the Suicide Behaviors Questionnaire-Revised (SBQ-R).

A number of behaviors have been used to define sexual risk behaviors as a syndemic condition. The most widely used of these criteria is condomless anal sex (N = `r sum(!is.na(measurement$CAS))`), followed by the number of partners (N = `r sum(!is.na(measurement$number_partners))`). The condomless anal sex criteria has been refined in three studies, two of them taking into account the type of partner (regular versus casual) and one, the serologic status of both partners.\
Other criteria such as condomless oral sex, condomless vaginal sex, STI diagnosis, engagement in sex work and sexual intercourse with an HIV positive partner have been used. Finally, one study used the Kalichman's Sexual Sensation Seeking Scale to assess the propensity of participants to engage in novel or risky sexual stimulation.

For violence `r .violence_prevalence$sum[["sexual"]]` studies looked at sexual violence, `r .violence_prevalence$sum[["physical"]]`, at physical violence and `r .violence_prevalence$sum[["psychological"]]` at psychological violence. One Indian study also took into account sexual and physical harassment by the police.

Several types of anxiety disorder were studied : generalized anxiety disorder (N = `r sum(!is.na(measurement$"generalized anxiety disorder"))`), social phobia (N = `r sum(!is.na(measurement$"social phobia"))`) and panic disorder (N = `r sum(!is.na(measurement$"panic disorder"))`). The type of anxiety disorder screened was not specified in `r sum(!is.na(measurement$"not specified"))` studies.\
`r length(unique(.anxiety_measurement$criteria))` different criteria were used to assess the presence of an anxiety disorder among the `r nodes$studies[nodes$label == "anxiety"]` studies considering anxiety as a syndemic condition. The most frequent criteria was a score equal or greater than 10 on the 7-items version of the Generalized Anxiety Disorder screening test (GAD-7 ; N = `r sum(!is.na(measurement$"GAD-7"))`).

SUD was screened through a clinical diagnosis based on the DSM-IV, mention of substance abuse in the electronic medical record or the participant thinking he should reduce his substance use. Several screening test were also used : the Drug Use Disorder Identification Test (DUDIT), the Texas Christian University Drug Screen (TCUDS), the Mini International Neuropsychiatric Interview (MINI) and the 10-items Drug Abuse Screening Test (DAST-10).

Outside of binge drinking and alcohol use disorder, which represents two distinct syndemic conditions, we aggregated three aspects of alcohol use studied in syndemic literature to form a generic "alcohol use" syndemic condition : alcohol consumption, alcohol use until intoxication and early alcohol use.

Finally, for discrimination, there were as many criteria as there were research papers studying this condition. Most of the studies only considered discrimination based on sexual orientation, except for two studies examining racial discrimination, one study examining HIV-based discrimination and one study which did not distinguish between discrimination based on sexuality, race/ethnicity or any other factor.

### Outcomes and their measurement

```{r outcomes_data, include=FALSE}

source("./scripts/create_outcomes_data.R", local = knitr::knit_global())

```

We identified `r ncol(.outcomes) - 1` different outcomes in our sample of quantitative studies. Sexual risk behaviors was the most frequently studied outcome, appearing in `r sum(!is.na(.outcomes$outcome_sexual_risk))` research papers. The second most studied outcome was HIV diagnosis (N = `r sum(!is.na(.outcomes$outcome_HIV))`), followed by STI diagnosis (N = `r sum(!is.na(.outcomes$outcome_STI))`), adherence to antiretroviral therapy (N = `r sum(!is.na(.outcomes$outcome_adherence))`), pre-exposure prophylaxis (PrEP) use (N = `r sum(!is.na(.outcomes$outcome_PrEP))`), healthcare use (N = `r sum(!is.na(.outcomes$outcome_healthcare))`) and the syndemic conditions themselves (N = `r sum(!is.na(.outcomes$outcome_syndemic))`). Less often, researchers also used viral load (N = `r sum(!is.na(.outcomes$outcome_VL))`), suicidality (N = `r sum(!is.na(.outcomes$outcome_suicide))`) or engaging in exchange sex (N = `r sum(!is.na(.outcomes$outcome_exchange_sex))`) as outcomes.

Every other outcome was only studied once : HIV screening[@Chandler2019b], physical activity[@Zhang2019], hypothalamic-pituitary-adrenal (HPA) axis dysregulation[@carrico2018], sexual violence[@semple2017], glycemic control[@Byg2016], engagement in HIV care[@Biello2016], engagement in chemsex[@friedman2014], having sexual intercourse with both men and women[@eaton2013] , help-seeking behaviors[@achterbergh2021] and substance use[@turpin2020].

There was a high degree of variability in the definition of what constitutes a sexual risk behaviors though the most frequently used proxy was, by far, condomless anal sex. Indeed, condomless anal sex was one of the criteria in `r .CAS_number` studies out of the `r sum(!is.na(.outcomes$outcome_sexual_risk))` with sexual risk behaviors as an outcome. While `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_CAS))` studies considered all form of condomless anal sex as a risk behavior, the other tried to refine the criteria. The most frequent of such refinement was to consider only serodiscordant condomless anal sex as a risk behavior (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_serodiscodant_CAS))`). Other options were to only consider condomless anal sex with casual partners (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_casual_CAS))`), condomless anal sex without PrEP or with a detectable viral load, if HIV-positive (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_PrEP_CAS)) + sum(!is.na(.outcomes_sex_risk_measurement$outcome_PrEP_VL_CAS))`), receptive condomless anal sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_receptive_CAS))`) or condomless anal sex at first sexual intercourse with the current male partner (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_current_CAS))`).

Used in conjunction with condomless anal sex, other proxies for sexual risk were number of partners (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_number_partners))`), substance use during sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_substance_sex))`), condomless vaginal sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_CVS))`), condomless oral sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_COS))`), sexual intercourse with female partners (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_female_sex))`) , engaging in group sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_non_disclosure))`) and non-disclosure of HIV serostatus before first sexual intercourse with current main partner (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_non_disclosure))`).

When HIV was used as an outcome, it was self-reported in nearly half of the studies (`r sum(.outcomes_measurement$measurement_outcome_HIV == "Self-report of HIV status", na.rm = T)` studies out of `r sum(!is.na(.outcomes$outcome_HIV))`) Similarly, for STI diagnosis, self-reporting was used in `r sum(.outcomes_measurement$measurement_outcome_STI_report == "Self-reported", na.rm = T)` papers out of `r sum(!is.na(.outcomes$outcome_STI))`).

### Interventions

We identified only `r .N_int` interventional studies guided by syndemic theory [@Chakrapani; @achterbergh2021]. Chakrapani et al. used a pre-test/post-test non-equivalent group design and the intervention consisted of motivational-interviews. The goal of this intervention was to reduce condomless anal intercourse by enhancing condom self-efficacy and addressing co-occurring syndemic conditions.

Achterbergh et al. conducted a randomized controlled trial and the intervention consisted of tailored feedback and help-seeking advice on mental health screening. The primary endpoint was to increase help-seeking behaviors ; the secondary endpoints were reducing sexual risk behaviors and STI incidence.

The intervention lasted for 12 months in both studies.

In Chakrapani et al. the intervention was successful in reducing the psychosocial conditions investigated (depression, alcohol use and internalised homophobia) and improving condom use. Furthermore, synergy, as measured by interaction on the additive and multiplicative scales, was present for depression and alcohol use as well as for depression and internalised homophobia on inconsistent condom use. Finally, mediation analysis revealed that the improvement in consistent condom use was due to an improvement in condom self-efficacy caused by a reduction in alcohol use and internalised homophobia.

On the other hand, the RCT conducted by Achterbergh et al. failed to affect the primary endpoint nor the two secondary endpoints.

### Summary of reviews

We found `r sum(summary_reviews$design == "Systematic review")` systematic reviews [@lewis2017; @woodward2017; @Lassiter2016] and 2 meta-analysis [@Rooney2018a; @Pantalone2020] in our review.

Woodward et al. aimed to identify a set of resilience resources among MSM burdened with minority stress and psychosocial condition in order to improve HIV prevention [@woodward2017]. They included `r summary_reviews$studies_included[summary_reviews$authors== "Woodward et al."]` in their reviews and identified 31 resilience resources, with social support and incomes as the two most frequently cited. Of note, most of these resources were associated with a lower HIV risk.

Lassiter et al. aimed to propose a framework to include religion and spirituality into HIV research with MSM [@Lassiter2016]. They found that religion and spirituality had mixed effect on syndemic conditions but could be more beneficial for MSM of color than for white MSM. Of note, they only included `r summary_reviews$studies_included[summary_reviews$authors== "Lassiter et al."]` studies as these two factors were very sparse in syndemic literature.

Lewis et al. sought to examine the HIV prevalence and associated risk behaviors among migrant and ethnic minority MSM in North America and Europe [@lewis2017]. They found high rates of HIV prevalence and associated risk factors and proposed that transnational migration could be part of a syndemic.

In their meta-analysis, Rooney et al. studied the syndemic conditions associated with sexual compulsivity among MSM and computed the mean effect size as well as whether this effect varied as a function of the type of syndemic conditions associated [@Rooney2018a]. They included `r summary_reviews$studies_included[summary_reviews$authors== "Rooney et al."]` papers and found that sexual compulsivity was significantly associated with 7 syndemic conditions (anxiety, depression, CSA, alcohol use, substance use, IPV and sexual risk behaviors). The two strongest associations were with depression and anxiety.

Finally, Pantalone et al. sought to have a better understanding of the state of interventions co-targeting interconnected syndemic conditions and HIV-related health behaviors of MSM [@Pantalone2020]. They included `r summary_reviews$studies_included[summary_reviews$authors== "Pantalone et al."]` studies and found a small significant positive effect of combined behavioral interventions to improve mental health, substance use, alcohol use and sexual risk behaviors, with significant heterogeneity. Interestingly, a greater number of sessions (at least 9) and individual rather than group interventions showed greater efficacy.

### Additional framework used in the studies

`r nrow(.framework)` studies used an additional framework alongside syndemic theory. The most widely used was resilience theory (N = 10), followed by minority stress model (N = 5), intersectionality (N=2), social-cognitive theory (N=2) and salutogenesis (N=1).

## How is the concept of interaction explored in syndemic research applied to MSM?

### Statistics used to show an interaction

```{r statistics_data, include=FALSE}

source("./scripts/create_statistics_data.R", local = knitr::knit_global())
```

It should be stated that, when we speak of interaction between syndemic conditions we don't necessarily mean "synergy", though synergy is indeed one of the interaction we consider. Drawing from the works of authors such as Tsai and Chakrapani, we consider three types of interaction : synergistically interacting epidemics, serially causal epidemics and mutually causal epidemics [@Tsai2018a; @Chakrapani2019]. As such, statistical methods such as mediation analysis, path analysis or structural equation modeling are also considered.

Our review revealed a high degree of variability in the statistic analyses. The most frequently used statistical method was to conduct regression analysis using a summation score of the syndemic conditions (N = `r sum(!is.na(summary_quantitative$score))`). Among studies employing this method, `r .score_data` did not use any other method to determine the degree of interaction between syndemic conditions.

In comparison, only `r .N_additivity` studies tried to evaluate departure from additivity on the additive and/or multiplicative scales, as recommended by Tsai [@tsai2015] to demonstrate synergy between syndemic conditions.

Furthermore, some studies sought to get a better understanding on the mechanism of interaction between the syndemic conditions and the outcome by using mediation analysis (N = `r sum(!is.na(summary_quantitative$mediation))`), moderation analysis (N = `r sum(!is.na(summary_quantitative$moderation))`), Structural Equation Modeling (N = `r sum(!is.na(summary_quantitative$SEM))`) or path analysis (N = `r sum(!is.na(summary_quantitative$path))`).

Moreover, `r .N_latent` studies used latent variables modeling such as Exploratory/Confirmatory Factor Analysis, Latent Profile Analysis, Latent Class Analysis or Latent Transition Analysis.

Using a different approach, `r sum(!is.na(summary_quantitative$network))` studies conducted by Lee and colleagues [@lee2020; @Lee] used network analysis, in which syndemic is conceptualized as a network and the syndemic conditions as interconnected nodes reinforcing each others.

Finally, other statistic analysis used to assess interaction were observed/expected ratio (N = `r sum(!is.na(summary_quantitative$O_E))`), cluster analysis (N = `r sum(!is.na(summary_quantitative$cluster))`) and point-biserial correlation matrix (N= `r sum(!is.na(summary_quantitative$biserial))`).

### Proposed mechanisms of interaction

In our sample of quantitative studies, we found only one study offering a mechanism of biological interaction @carrico2018 and one study offering a mechanism of bio-social interaction [@Klein2011].

Carrico et al. showed that the combined effects of HIV infection and methamphetamine use were detrimental for the functioning of the Hypothalamic-Pituitary-Adrenal axis which is thought to play an important part in the reinforcing effect of stimulants.

Klein proposed that attitudes towards condom was one of the key factor contributing to condomless sex and, subsequently, HIV infection. In their study, attitudes toward condoms was predicted by low self-esteem, as condom use represents a self-protecting mechanism, as well as by sexual preferences, substance use, race and education. Furthermore, childhood emotional neglect had a negative influence on self esteem thus indirectly contributing to sexual risk taking.

Qualitative studies proposed mechanisms of bio-social interaction more often than quantitative studies as we identified `r sum(!is.na(summary_qualitative$bio_social_interaction))` papers with such hypothesis of interaction. On the other hand, we did not find any qualitative paper exploring biological interaction.

Stigma and structural inequalities were indicated by the 5 papers as the root causes of syndemic conditions and HIV risk behaviors. Maionara et al. showed that economic dependency and fear of violence may maintain YBMSM in dysfunctional relationships plagued by IPV and substance abuse which increase their odds of incarceration and HIV infection [@Maiorana]. Furthermore, methamphetamine is used as a coping strategy while simultaneously representing a potential factor in acquiring or transmitting HIV to others, notably during sex parties.\
This finding of substance use as a coping mechanism was supported by the study by Pollard et al. who found that maladaptive coping strategies to minority stress and the performative resistant space of the gay scene contributed to chemsex use [@Pollard2018a].

For Black MSM, other sources of structural violence such as racial neighborhood segregation contributed, alongside poverty, unemployment, violence, unstable housing, incarceration and poor healthcare access to expectations surrounding masculinity which may conflict with the sexual orientation of BMSM, leading to internalised homophobia [@Quinn]. Internalised homophobia, in turn, may lead to poor mental health as well as to substance abuse and sexual risk taking as coping mechanisms. Moreover, the use of PrEP, perceived as a "gay pill" may negatively impact the use of PrEP in this community. Lyons et al. also showed that the marginalization of YMSM within their school, communities of origin and families resulted in a lack of gay-specific HIV prevention education, role-model and productive goal-related activities [@Lyons2013]. The HIV-positive MSM youth in this study linked these factors to their HIV acquisition. Finally, [@Chakrapani2019a] showed the role of family support and communities of MSM to serve as resilience resource to counteract the negative effects of stigma and syndemic conditions.

## What were the key findings of these studies?

In nearly all of the quantitative studies, the syndemic conditions chosen by the authors were found to be associated to the outcomes of interest.

However, some interesting discrepancies in the findings were found. On the topic of synergy, results were conflicting between studies. Chakrapani and colleagues found support for the three models of interacting epidemics (synergistically interacting epidemics, serially causal epidemics and mutually causal epidemics) with the strongest support for the model of synergistically interacting epidemics [@Chakrapani2019]. These findings were supported by other studies who found synergy when using significant product terms in the additive and/or multiplicative scales [@Chakrapani; @chandler2020; @Chandler2019; @tomori2018]. However, other studies found no synergy between the syndemic conditions using the same statistical approach [@Batchelder2019; @shuper2020; @Ferlatte2018a].

Conflicting results were also found for MSMW. Brandstrom and Pachankis found no clustering of syndemic conditions for MSMW [@Branstrom2018] while Mustanski et al. found a stronger clustering of syndemic conditions in YMSMW compared to YMSM[@Mustanski2014]. Furthermore, Dyer et al. found more STI in the Black MSM compared to the Black MSMW [@dyer2020] and Ferlatte found more syndemic conditions in the MSM compared to the MSM [@ferlatte2018], mainly driven by party drug use and treatment for depression or anxiety. Finally, in a sample of Latino MSMW, having two syndemic conditions predicted receptive condomless anal sex and STI diagnosis but not insertive nor vaginal condomless sex [@Munoz-Laboy2018].

While condomless anal sex is often used as a proxy for the risk of acquiring HIV or another STI, the findings of some studies tend to show some limitation of this proxy. For example, Moeller et al. found an association between syndemic conditions and condomless anal sex but not with HIV diagnosis [@moeller]. A similar pattern was found in a longitudinal study that found an association between syndemic conditions and sexual risk behaviors but not with HIV or STI incidence [@Mustanski2017]. On the other hand, some authors found an association between syndemic conditions and HIV but not with condom use [@chuang2018].

Alcohol misuse was inconsistently linked to risk taking. Card et al found no association between the AUDIT score and sexual risk behaviors [@Card2018] while another study using latent class analysis found that the class with alcohol misuse had a lower level of risk behaviors [@scheer2021]. Nonetheless, alcohol misuse was found to be associated with PrEP non-adherence [@shuper2020].

In a longitudinal study [@Mustanski2017] Black MSM were found to have the highest HIV incidence while having a lower burden of syndemic condition than White MSM, raising the question of the need to focus on syndemic conditions specific to Black MSM.

Social support was shown to moderate the effect of syndemic conditions on viral load, providing a significant protective effect [@friedman2016]. Moreover, a paper using Latent Transition Analysis to evaluate the impact of syndemic conditions on substance use showed that Black MSM with more social support were more likely to stay in the low-risk class or transitioning from the high risk class to the low-risk class [@turpin2020]. However, Chakrapani et al. found no moderation of social support on the relationship between syndemic conditions and risk taking [@Chakrapani2017].

On the topic of resilience resources, Zhang et al didn't find evidence of a moderating effect of resilience on the relationship between syndemic conditions and physical activity [@Zhang2019]. O'Leary et al. didn't find a buffering effect of resilience factors on the relationship between syndemic conditions and sexual risk behaviors, though optimism and education buffered the relationship between syndemic conditions and self-reported HIV status [@OLeary2014a]. On the other hand, Kurtz et al found that, among MSM living with HIV, serosorting was positively associated with higher levels of two resilience factors : coping self-efficacy and positive coping skills [@Kurtz2012]. Similarly, Hart et al. found that psychosocial strengths were associated with lesser odds of engaging in condomless anal sex, despite the presence of syndemic conditions [@Hart2017].