### 2021-10-19 _bookdown_files/06-conclusion.Rmd
Conclusion

# Strengths and limitations

This review is the first Scoping Review to map the current knowledge in Syndemic Theory applied to MSM. We charted and presented detailed data regarding the way research has been done. As such, we offer a precious insight and actionable recommendations for the future of the field. Our search strategy of electronic databases was comprehensive and developed with the help of the director of health sciences library of our institution. We combined this database search with a secondary hand-search of references lists of the included studies. Moreover, we conducted a fully transparent and reproducible review. Other researchers may freely reuse our data and scripts to conduct their own work. To this end, we created a SQL database from our data. Other researchers may thus perform complex queries on our database of included studies to quickly identify studies relevant for their work (e.g. every study using the CESD to measure depression on Black MSM between 2014 and 2019).

However, even though we took several steps to ensure the comprehensiveness of our search strategy, we may still have missed some relevant literature. Furthermore, one of our inclusion criteria was the centrality of syndemic theory to the screened paper. Even if two reviewers screened a part of the studies independently, this criteria may be somewhat arbitrary and other researchers may have chosen to include papers we rejected or the contrary. Finally, usual limitations of a scoping review apply such as the lack of risk of bias and strength of evidence assessment.

# Conclusion and recommendations

Our review of syndemic research applied to MSM revealed some important limitations that must be addressed in future work. The field of syndemic research would gain to study more MSM subpopulations, in more diverse geographical settings, especially in the Global South. Furthermore, promising methods such as spatial epidemiology or network analysis may help to advance the field while methods such as summary count should be abandoned. Besides, more qualitative and mixed data may enhance our understanding of the complex interactions of a syndemic on the living experiences of marginalized people suffering from it. Moreover, outcomes should be less focused on sexual health and start to explore other potential impact of a syndemic on health. Finally, conditions that constitute a syndemic should be more precisely defined and the measurement of said conditions should be more standardized while taking into account the difficulties of transposing some measures in different sociocultural contexts.

Though we took a critic stance on the current literature, we still consider syndemic to be an useful framework to study the global health of marginalized populations and the current research still offered valuable insights on the health of MSM. It is up to scholars in the field to continue to push research forward in order to improve the health of marginalized populations worldwide.