### 2022-01-13 _bookdown_files/index.Rmd

---
title: "Mapping the current knowledge in syndemic research applied to Men who have Sex with Men : a Scoping Review"
author: 
- "Maxence R. Ouafik^1^, Laetitia Buret^2^, Jean-Luc Belche^3^, Beatrice Scholtes^4^"
- "^1^ Research Unit of Primary Care and Health, General Practice Department University of Liège, Liège, Belgium ; Quartier Hôpital B23 Avenue Hippocrate 13, 4000, Liège <maxence.ouafik@uliege.be> ; Quartier Hôpital B23 Avenue Hippocrate 13 4000 Liège ![OrcID](./Files/Images/OrcID.png) https://orcid.org/0000-0002-9795-5721"
- "^2^ <laetitia.buret@uliege.be> ![OrcID](./Files/Images/OrcID.png) https://orcid.org/0000-0001-6039-9824"
- "^3^  <jlbelche@uliege.be> ![OrcID](./Files/Images/OrcID.png) https://orcid.org/0000-0001-8807-0473"
- "^4^  <beatrice.scholtes@uliege.be> ![OrcID](./Files/Images/OrcID.png) https://orcid.org/0000-0001-5274-822X"
- "*corresponding author: maxence.ouafik@uliege.be https://orcid.org/0000-0002-9795-5721"

link-citations: true
csl: ["./_bookdown_files/social-science-and-medicine.csl"]
bibliography: ["./_bookdown_files/references.bib"]
rmd_files: ("index.Rmd", "01-abstract.Rmd", "02-introduction.Rmd", "03-methodology.Rmd", "04-results.Rmd", "05-discussion.Rmd", "06-conclusion.Rmd", "07-references.Rmd", "08-appendix.Rmd")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(bookdown)
library(officedown)
library(officer)
```