### 2022-01-23 _bookdown_files/05-discussion.Rmd

---
bibliography: ./_bookdown_files/references.bib
---

# Discussion

The objective of this scoping review was to map the current knowledge on syndemic research applied to MSM. To this end, we analysed 115 articles and found that the literature was both too homogeneous in some aspects while being not focused enough in other aspects. As we will discuss in more depth below, the location, design, subpopulation and outcomes lacked diversity. On the other hand, the syndemic conditions and ways of measurement were not standardised enough to ensure the robustness, reproducibility and comparability of the findings. Furthermore a substantial part of the literature doesn't provide empirical data to support one of the core tenets of syndemic theory - namely disease interaction- as has been pointed out in a previous review [@tsai2015].

## Location

The location of study is an aspect of syndemic literature suffering from too much homogeneity. Indeed `r percent(.North_America_studies/.N_QualQuant)` of studies are conducted in North America alone while all studies conducted in the Global South represents only `r percent(.GlobalSouth_studies/.N_QualQuant)` of our sample.

Three issues arise from this. First, under-representation of the Global South compromises our understanding of the health of MSM living in these countries. Furthermore, as pointed out by @weaver2020, the screening tool used to assess the presence of a syndemic condition in the Western context may not be efficient in another sociocultural setting, especially in regard to mental health. As syndemics are driven by the sociocultural forces of a given place at a given time, these potential biases need to be addressed in order to conduct meaningful research outside Europe and North America.

Moreover, even for studies conducted in Europe and North America, a large portion of the studies conflate data from different cities, counties, states or even countries in order to obtain a sample big enough to conduct statistical analyses. However, geographical variations in syndemic burden are likely to be missed. Migration of MSM to larger cities is a well-known phenomenon and thought to be an important part of syndemic production among MSM as well as a source of resilience [@stall2007; @Bruce2011]. Differences in the density of gay social venues and peer organizations as well as exposure to discrimination and violence or an easier access to substances need to be taken into account when studying a syndemic. As @shrestha2020 suggested, syndemic research needs to incorporate data from spatial epidemiology to better understand the interactions of socioeconomic forces in the physical space and how these forces can influence the health of MSM.

## Design

The design of syndemic studies is another element suffering from a lack of diversity. The field is dominated by quantitative studies as qualitative papers represent only `r percent(.N_qual/nrow(overview_of_studies))` of our sample. Moreover `r percent(.N_cross_sec/nrow(summary_quantitative))` of the quantitative studies were cross-sectional. As vulnerability to syndemic conditions is suspected to arise from early adverse experience @stall2007 longitudinal studies running over a long period of time are probably needed to study syndemic production among MSM.

Moreover, as shown in our result, qualitative papers are more prone to discuss the potential interactions between syndemic conditions. Some concepts and experiences needed to understand syndemic production in MSM can only be studied through a qualitative approach. For example, the complex narratives surrounding drug use and its link to loneliness, marginalization, personal affirmation, resistance and lack of meaningful emotional connections @Pollard2018a would be hard to measure and model using quantitative data. As Tsai previously pointed out, the inevitable simplification needed to estimate statistical models may lead to an over-simplification of diseases dynamics @Tsai2018a . Mixed method are needed to complement robust statistical estimation with insights on sociocultural mechanisms and on the lived experiences of people facing this clustering of adverse conditions. Unfortunately, our review only identified `r max(year_design$cumulative[year_design$design == "Mixed"])` studies employing such mixed methodology. Future research would benefit from generating and analysing qualitative data to support and enrich their quantitative findings.

Finally, we identified `r .N_int` interventional studies with mixed results. While co-targeting interconnected syndemic conditions is supported by a meta-analysis @Pantalone2020, a better understanding of how syndemics operate may be needed before we are able to offer efficient syndemic-based interventions.

## Subpopulations

Studying the health of MSM poses a challenge regarding the diversity of this population. While nearly half of the studies included in our sample studied a subpopulation of MSM, the number of MSM subpopulation is such that most MSM subpopulation remains under-served. For example, Young Black MSM are specifically studied in `r .N_YBMSM` studies while representing 52% of the new HIV diagnosis among MSM aged 13-24 years in 2018 @centerfordiseasecontrolandprevention2020.

This discrepancy between the number of studies focusing on a subpopulation and the very high prevalence of some syndemic condition such as HIV infection is even more concerning when considering that dual-minority identity may expose a person to different stressors and outcomes. For example, some studies found that Black MSM had a higher rate of HIV seroconversion than White MSM while being exposed to fewer syndemic conditions @Mustanski2017 . These findings are supported by qualitative data suggesting that intersectional theory offers the context needed to understand syndemics among Black MSM which differ from syndemics among White MSM @Quinn. Furthermore, the findings from Bulled and colleagues suggest that structural disadvantage may have more weight on the lives of Black MSM than individual factors such as substance use @bulled2021

MSMW are another example of an under-served population with conflicting results. In a cross-sectional study using syndemic sum count, Latino MSMW with two or more syndemic conditions had a 7.09 Odds Ratio to engage in receptive condomless anal sex but no statistically significant difference was found for insertive condomless anal sex or condomless vaginal sex @Munoz-Laboy2018. On the other hand, in another study the number of syndemic conditions was associated with HIV diagnosis for MSM but not for MSW or MSMW @Branstrom2018. Among young MSM and MSMW, one study found more syndemic conditions and a greater association between them for MSMW compared to MSM @Mustanski2014. On the contrary, Ferlatte et al. found that MSM were more likely than MSMW to suffer from two or more syndemic conditions @ferlatte2018. These conflicting findings may be partly due to the difficulty to define and operate bisexuality in research @swan2018, differences between age groups or differences in syndemic mechanism for MSMW compared to MSM. Outside the field of syndemic literature, studies suggest that bisexual men suffer from a similar or higher burden of mental health conditions compared to gay men [@salway2019; @Chaudhry2019; @ross2018] and literature specific to bisexual men remains scarce. In the future, syndemic theory should thus seek to study the specificity of this population. First, there's a need to better conceptualize and define bisexuality in the field - should we still use solely behavior-based definition (as in MSMW) or should we incorporate a self-definition of one's sexual orientation ? Then, researchers should aim to disaggregate data concerning MSMW/bisexual men from data concerning MSM/gay men whenever possible. Lastly, more research focusing on MSMW/bisexual men is needed.

Finally transgender MSM were included in only one study in our sample. As stated in our results, this is concerning, especially when every other study excludes transgender men by design. As such, this population of MSM is the one we know the least about. The study from Reisner and colleagues suggested that transgender MSM who had socially affirmed their gender were exposed to similar risk patterns as cisgender MSM @Reisner2016a. The health of transgender men remains greatly understudied and the few available studies suggest high rates of adverse conditions such as violence, notably sexual violence @Testa2012, suicide attempts @Haas2014, chlamydial and gonoccocal infections @Pitasi2019, intimate partner violence [@peitzmeier2020] and HIV infection @Becasen2019. Researchers should be encouraged to include transgender men in their studies and to conduct research focused on transgender MSM. Indeed, syndemic theory represents a promising framework for research on transgender men as its premises of social adversities producing and concentrating diseases in a population allows us to explore the health of transgender men without pathologizing them.

## Syndemic conditions

There is a lack of consensus in the choice of syndemic conditions and the way they should be measured. We identified `r nrow(nodes)` different conditions in our quantitative sample, `r nrow(nodes %>% filter(studies == 1))` of them were only studied once. Despite this diversity of syndemic conditions, the impact of the seminal study by Stall et al.[@Stall2003] is striking. The conditions studied in that paper (depression, IPV, CSA and polysubstance use) are among the five most studied conditions in the field. Furthermore, `r percent(nrow(.Stall_conditions)/nrow(summary_quantitative))` of the quantitative studies included in our review consider at least one of the four conditions studied by Stall and colleagues as part of a syndemic.

Some of these conditions are past event (e.g. history of trauma or childhood sexual abuse) while most were current conditions. Given the cross-sectional nature of `r percent(.N_cross_sec/nrow(summary_quantitative))` of the quantitative papers, we cannot disentangle how past experiences may influence the emergence of a syndemic.

Furthermore, studies diverge on what represents a syndemic condition and what may lead to the emergence of a syndemic. For example, childhood sexual abuse is mostly seen as a syndemic condition per se but may very well be a risk factor to the emergence of a syndemic, as hypothesized by some authors [@Herrick2013; @leblanc2021]. The same question may be asked for discrimination, trauma, impulsivity, bullying, poor social support, loneliness, childhood abuse, poor healthcare access, etc. Qualitative and longitudinal studies are needed to distinguish between syndemic conditions and pathways to the emergence of a syndemic. Conflating the two in cross-sectional studies does very little to advance our understanding of syndemic theory or to implement public health interventions.

Furthermore, even when the same condition was studied in multiple papers, the differences in measurement methods impedes the reproducibility of the results. For example, a participant with a CESD score of 16, who uses marijuana on a regular basis and cocaine once in a while and who is a victim of sexual intimate partner violence. This man would have been classified as depressed in `r percent(sum(.depression_measurement$depression_cutoff == 16 & .depression_measurement$depression_scale_criteria == "CESD", na.rm = T)/sum(.depression_measurement$depression_scale_criteria == "CESD"))` of studies using the CESD, a polysubstance user in `r percent(sum(.polysubstance_measurement$polysubstance_number == "More than one category of substance use" & .polysubstance_measurement$polysubstance_marijuana == "Marijuana included")/nrow(.polysubstance_measurement))` of studies considering polysubstance use and as a victim of intimate partner violence in `r .IPV_prevalence$perc[["sexual IPV"]]` of studies taking IPV into account. In other words, the same person could have been classified as having 0, 1, 2 or even 3 syndemic conditions, depending on the study design. Considering that most studies use a summation score of the number of syndemic conditions to conduct regression analysis, it seems evident that the results would greatly vary according to the way the syndemic conditions are measured.

It is beyond the scope of this paper to discuss in detail which syndemic conditions should be retained for future research and how to best measure them, but without a greater homogeneity in the field, the confidence and reproducibility of the results will remain an issue.

## Outcomes

Taken together, `r percent(.HIV_number/nrow(summary_quantitative))` of the quantitative studies in our sample had an HIV-related outcome, whether in the form of sexual risk behaviors, HIV diagnosis, HIV screening, adherence to antiretroviral therapy, PrEP use, viral load or engagement in HIV care.

This omnipresence of HIV in syndemic literature applied to MSM is understandable. The first description of a syndemic, the SAVA syndemic, incorporated AIDS as part of a syndemic @singer1996. Then, a few years later, the first paper on syndemic in the MSM population had HIV-related outcomes, in the form of HIV serostatus and CAS @Stall2003. Furthermore MSM were the first and most severely affected population when the AIDS epidemic started @decock2012. Nowadays, male-to-male sexual contact still represents nearly 70% of the new HIV cases in the USA @centerfordiseasecontrolandprevention2020. While we can't deny the importance of the HIV pandemic, notably for MSM, we would argue that HIV is not the only issue in need of attention.

Indeed, MSM are also disproportionately affected by, amongst others, suicide, anxiety, depression and substance use[@ross2018; @luo2017; @Medley2016]. Moreover, with the advances in antiretroviral therapy and the subsequent decline in HIV-related mortality, a Canadian study showed that gay men were more likely to die from suicide than from HIV in 2011 @hottes2015. Even if the estimates Hottes' paper overestimated the mortality due to suicide among MSM, suicide would still be a leading cause of death in this population while being studied as an outcome in only `r sum(!is.na(.outcomes$outcome_suicide))` papers in our sample.

Furthermore, when studying HIV as an outcome, we would argue that using condomless anal sex as a proxy should be reviewed and better proxies sought. Indeed, condomless anal sex was one of the criteria in `r .CAS_number` studies out of the `r sum(!is.na(.outcomes$outcome_sexual_risk))` with sexual risk behaviors as an outcome and `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_CAS))` studies considered all form of condomless anal sex as a risk behavior. However, without refinement, this outcome fails to capture the changes in HIV prevention strategies such as PrEP or Treatment as Prevention (TasP). Only `r english(sum(!is.na(.outcomes_sex_risk_measurement$outcome_PrEP_VL_CAS)))` studies in our sample took PrEP use and an undetectable viral load into account when considering condomless anal sex as a risk factor for HIV acquisition. When the first paper on syndemic theory applied to MSM was published, engaging in condomless anal sex would have led to a high risk of acquiring HIV. Nowadays, people could be on PrEP, forego the use of a condom and still be at a very low risk of acquiring HIV @calabrese2017 and our methodologies in syndemic research should reflect this evolution. In addition, qualitative research gives an insight into risk behaviors and offers a more nuanced view on this topic than a binary variable of consistent condom use for anal intercourse. One such paper published in 2017 showed that risk practices fell into different subjectivities such as active and consistent pursuit of condomless sex, lack of assertiveness to a partner's initiative of condomless anal sex and combination of episodic risk practices and reduction strategies @Adam2017. Besides, some studies showed that participants with a greater number of syndemic conditions would engage more frequently in condomless anal sex while not being more likely to test positive for HIV [@Mustanski2017; @Pitpitan2016]. On the other hand, a paper from 2018 found that having a greater number of syndemic conditions was associated to self-reported HIV but not to condomless anal sex @chuang2018. Taken together, we thus postulate that condomless anal sex should at least be replaced by "condomless anal sex without consistent PrEP use or with a detectable viral load" when an investigator seek to use sexual risk behaviors as an outcome.

In summary, we recommend that future research should focus on other aspects of MSM's health than HIV infection and, when studying HIV as an outcome to prefer laboratory-based outcomes or more refined sexual risk variables than condomless anal sex.

## Statistics and interaction

One of the tenets of syndemic theory is that there exists some form of biological, social and/or psychological interaction between the diseases @mendenhall2020. As others have already pointed out, the extent to which syndemic literature had proved this core principle remains very limited @tsai2015. Indeed, `r percent(.score_data/nrow(summary_quantitative))` of our quantitative sample of studies used solely a summation score of the syndemic conditions to conduct regression analysis. Two issues arise from this statistical approach. First, summation score are additive by essence, thus unsuitable to demonstrate an interaction @tsai2015. Second, two psychometric assumptions are made by such a model : (1) the unidimensionality of the construct and (2) equal factor weighting @Halkitis2013a. Put simply, using a summation score would imply that (1) every condition forms a single construct and that (2) every condition contributes equally to said construct (which would mean for example that feeling depressed during the past 14 days and having an history of childhood sexual abuse would contribute equally to a syndemic). The unidimensionality aspect of the construct has received some empirical support by Mustanski et al. as well as by Starks et al. using respectively Structural Equation Modeling and Latent Class Analysis [@Mustanski2014; @Starks2014]. On the contrary, Leblanc et al. failed to construct a single latent syndemic variable in their analysis [@leblanc2021]. Moreover constraining factor loadings to be equal resulted in worse fit in the studies by Mustanski and Starks.

In short, the result of our review suggests that as much as `r percent(.score_data/nrow(summary_quantitative))` of the quantitative literature on syndemic theory applied to MSM doesn't empirically support the presence of a syndemic and employs questionable statistical assumptions. Those studies are more an exploration of psychosocial risk factors associated to HIV acquisition or progression than the holistic framework syndemic is supposed to be.

As mentioned in the results, two statistically distinct models may be envisioned apart from synergistically interacting epidemics : mutually causal epidemics and serially causal epidemics [@Tsai2018a]. As such, mediation analysis, path analysis or structural equation modeling may prove useful to assess an interaction.

More recently, a new statistical approach inspired from research in psychopathology has been used to model syndemics : network analysis [@Lee; @lee2020]. The core principle of this approach is that mental disorders arise from causal interactions between symptoms in a network @borsboom2017. Applied to syndemic theory, a syndemic would then be the network and the various conditions would then be the nodes of the network. Furthermore, one of the benefits of this approach resides in the theoretical possibility to identify influential nodes in a network : nodes that play a significant part in the activation of the network and that could be the prime target of an intervention in order to deactivate the network @robinaugh2016. In short, in addition to being statistically sound and coherent with the model of mutually causal epidemics, network analysis could offer actionable data to improve the health of MSM by identifying which conditions should be prioritized for interventions [@Tsai2018a].

We recommend that future research should abandon the summation score approach and attempt to clarify which model of interaction receives the most empirical support in order to implement effective public health interventions. Indeed as reminded by Chakrapani et al. the three models have different programmatic implications @Chakrapani2019. In the case of synergistically interacting epidemics, a single intervention may yield greater health improvement than if no interaction was present. For mutually causal epidemics, multicomponent interventions addressing the various epidemics should be implemented in order to produce an improvement. Finally, for serially causal epidemics, intervening at the root cause of a syndemic may prevent the development of a syndemic cascade.

## Additional frameworks

Resilience theory was the most frequently used framework to compliment syndemic theory.

Social support was shown to moderate the effect of syndemic conditions on viral load, providing a significant protective effect [@friedman2016]. Moreover, a paper using Latent Transition Analysis to evaluate the impact of syndemic conditions on substance use showed that Black MSM with more social support were more likely to stay in the low-risk class or transitioning from the high risk class to the low-risk class [@turpin2020]. However, Chakrapani et al. found no moderation of social support on the relationship between syndemic conditions and risk taking [@Chakrapani2017].

On the topic of resilience resources, Zhang et al did not find evidence of a moderating effect of resilience on the relationship between syndemic conditions and physical activity [@Zhang2019]. O'Leary et al. did not find a buffering effect of resilience factors on the relationship between syndemic conditions and sexual risk behaviors, though optimism and education buffered the relationship between syndemic conditions and self-reported HIV status [@OLeary2014a]. On the other hand, Kurtz et al found that, among MSM living with HIV, serosorting was positively associated with higher levels of two resilience factors : coping self-efficacy and positive coping skills [@Kurtz2012]. Similarly, Hart et al. found that psychosocial strengths were associated with lower likelihood of engaging in condomless anal sex, despite the presence of syndemic conditions [@Hart2017].

In qualitative studies, resilience was also found to exert a protective effect against syndemic conditions [@Chakrapani2019a; @Reed2016; @adam2018]. Furthermore, the review by Woodwards et al. found a protective effect of most resilience resources identified against HIV @woodward2017.

We recommend that future work continues to explore the impact of resilience in partially counteracting syndemics to develop public health interventions that aim to foster resilience resources among MSM. Moreover, as pointed out by Namer and Razum, focusing on risk rather than resilience and survivorship poses the threat of alienating minority groups rather than prioritizing them [@namer2021].

Minority stress represents another interesting framework as it allows to explain how social stigma can get under the skin of minorized populations @Meyer2003.

Finally, although intersectionality was only used by two studies [@ferlatte2018; @Quinn] we believe the field would benefit from this framework. Indeed as shown in our results, there is some discrepancy in the results for MSM's subpopulation, especially for MSMW/bisexual men, Black MSM and Latino MSM. We believe that the inattention to systemic inequalities and structural disadvantage in the field may partly explain these discrepancies. In fact, only `r percent((summary_quantitative %>% filter(!is.na(unstable_housing) | !is.na(discrimination) | !is.na(incarceration) | !is.na(poverty)| !is.na (unemployment) | !is.na(healthcare_access)) %>% nrow())/nrow(summary_quantitative))` of the quantitative studies examine a structural condition such as discrimination, poverty or access to healthcare.

Some scholars are skeptical on the usefulness of syndemic theory to intersectionality and question the possibility of a conversation between the two fields @sangaramoorthy2021. As pertinent as the authors commentary is, we do believe in the possibility of making the two field converse. As Quinn pointed out in her response, incorporating an intersectional framework into syndemic research may address issues of power, oppression and structural violence, neglected thus far in the field @quinn2021. Doing so, we may attend to some gap in the literature and disentangle the complex web of stigma and structural violence faced by sexual minority men, especially MSM situated at the intersection of multiple form of systemic discrimination such as racism, homophobia, biphobia, transphobia, classism, ableism or stigma against sex workers or people living with HIV.