### 2022-04-22 _bookdown_files/04-results.Rmd

---
bibliography: ./_bookdown_files/references.bib
---

```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

```{r packages_data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

#Load required packages:
library(tidyverse)
library(readr)
library(scales)
library(ggplot2)
library(ggpubr)
library(sunburstR)
library(d3r)
library(english)

#Load the data from the.csv files
source("./scripts/read_data.R", local = knitr::knit_global())

```

# Results

This section summarises the findings relevant to our research questions. Table \@ref(tab:RefTabQuant), \@ref(tab:RefTabQual) and \@ref(tab:RefTabRev) in the supplementary material, summarise every reference identified and the variables extracted to answer our research questions for quantitative studies, qualitative studies and reviews, respectively.

## Literature Search

The electronic search identified 840 references and our hand search of reference lists resulted in the addition of the seminal study by Stall and colleague [@Stall2003]. After removing duplicates, 305 records were screened for inclusion.

After screening the title and abstract, 124 records were excluded. The two main reasons for the exclusion at this stage were that MSM were not the main study population (n=49) and that syndemic was not the main focus of the paper (n=37). Thirty-eight studies were excluded because the type of publication did not meet our inclusion criteria.

The full texts of the 181 remaining references were obtained and read; Sixty-six were excluded after this phase. During the previous step, we had decided to tentatively include papers in which the sample was not entirely comprised of MSM. After reviewing the full paper, we decided to exclude studies if the MSM data were aggregated with data from another population (n=11), most commonly transgender women. The rationale for this decision was to keep the focus clearly on MSM.

The PRISMA flowchart generated by DistillerSR can be found in Figure \@ref(fig:PRISMA)

```{r PRISMA, fig.cap= "PRISMA flowchart of the search strategy", fig.height=7.5, fig.id="PRISMA", fig.width=7, fig.align='center'}
knitr::include_graphics("Files/images/PRISMA.png")

```

## How are studies concerning Syndemic Theory applied to MSM conducted?

### General characteristics of included studies

```{r general_characteristics_data_preparation, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

#Source the scripts used to manipulate the data from the.csv files
source("./scripts/create_year_design.R", local = knitr::knit_global())
source("./scripts/create_location_data.R", local = knitr::knit_global())
source("./scripts/create_sample.R", local = knitr::knit_global())

#Create the plot 
source("./scripts/create_plot_year.R", local = knitr::knit_global())

```

The publication date of the studies ranges from `r min(year_design$year)` to `r max(year_design$year)` with a marked increase in the number of papers published annually from 2013 onward as illustrated in Figure \@ref(fig:PlotYear). There is also an increase in the diversity of the studies, both in the study design used and the continents where the studies were conducted.

```{r plot_year, fig.cap="Plot of the cumulative number of studies published yearly", fig.height=7, fig.id="PlotYear", fig.width=5.5, fig.align='center'}

knitr::include_graphics("./Output/Plots/Plot_Year.png")

```

However, despite the increasing diversity, the vast majority of the studies were conducted in North America (N = `r .North_America_studies`), with a large predominance of studies in the USA (N= `r .USA_studies`). The majority of the remaining studies were then located in Asia (N = `r .Asia_studies`) or in Europe (N = `r .Europe_studies`). South America (N = `r .South_America_studies`) and Africa (N = `r .Africa_studies`) are the least represented continents.

In terms of design, a similar pattern appears, with most of the studies employed a cross-sectional design (N = `r .N_cross_sec`). The number of longitudinal studies has grown steadily since 2014 and represents `r percent(.N_long/nrow(overview_of_studies))` of the papers. The duration of longitudinal studies ranged from `r min(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` to `r max(summary_quantitative$study_duration[summary_quantitative$design == "Longitudinal"], na.rm = T)` months.

Compared to quantitative studies, qualitative studies are under-represented, with only `r .N_qual` papers. In terms of analysis method, the most common is thematic (content) analysis (N= `r sum(!is.na(summary_qualitative$thematic_content)) + sum(!is.na(summary_qualitative$thematic))`). The rest of the studies consists of framework analysis (N = `r sum(!is.na(summary_qualitative$framework))`), grounded theory (N = `r sum(!is.na(summary_qualitative$grounded))`), analytic induction (N = `r sum(!is.na(summary_qualitative$analytic_induction))`) and constant comparative analysis (N = `r sum(!is.na(summary_qualitative$comparative))`).

We identified 2 papers employing mixed methods design [@buttram2015; @Halkitis2012]; Buttram et al. used a cross-sectional design for the quantitative part and in-depth interviews analysed through grounded theory for the qualitative part while Halkitis et al. used a cross-sectional quantitative survey and in-depth discovery interviews.

The first systematic reviews and meta-analyses were published in `r min(year_design$year[year_design$design == "Systematic reviews, meta-analyses"])`, while the first interventional studies were published in `r min(year_design$year[year_design$design == "Interventional"])`.

Finally, the median sample size of the studies is `r median(sample$sample_size)` (range: `r min(sample$sample_size)` - `r number(max(sample$sample_size))`) and the mean age of the participants, when reported, ranges from `r min(sample$age, na.rm = T)` to `r round(max(sample$age, na.rm = T), digits = 0)` years.

### Subpopulations studied

```{r Subpopulation_data_prep, message=FALSE, warning=FALSE, include=FALSE}
source("./scripts/create_population_data.R", local = knitr::knit_global())
source("./scripts/create_population_plot.R", local = knitr::knit_global())

```

Less than half of the studies (N = `r sum(!is.na(population$subpopulation))` ) focus on a subpopulation of MSM; of those, we identified seven types of MSM subpopulations: (a) young MSM; (b) older MSM; (c) MSM from a racial/ethnic minority; (d) MSM living with HIV; (e) MSM engaged in sex work; (f) Men who have Sex with Men and Women (MSMW); and (g) transgender MSM.\
Among studies focusing on a specific age group, `r sum(population$age == "Young adult")` studies focus on young adults, `r english(sum(population$age == "Older men"))` studies focus on older MSM [@zepf2020; @Halkitis2012] and `r english(sum(population$age == "Adolescent"))` study focuses on adolescents [@Perry].\
Black MSM (BMSM) are the most studied MSM from a racial/ethnic minority (N= `r sum(population$race_ethnicity == "Black")`) but studies focusing on them still represent only `r percent(sum(population$race_ethnicity == "Black")/104)` of our sample.\
`r english(sum(population$race_ethnicity == "Latino"))` studies focus on Latino MSM (LMSM) and `r english(sum(population$race_ethnicity == "Latino & Black"))` study focus on both BMSM and LMSM [@cassels2020].

Concerning the serologic status of MSM in syndemic literature, `r sum(population$serologic_status == "HIV-positive")` studies are entirely comprised of a sample of MSM living with HIV. Notably, we didn't identify any studies focusing on racial minority MSM living with HIV.\
Finally, the least represented subpopulation of MSM in this review are MSM engaged in sex work (N = `r sum(population$sex_work == "Sex workers")`), MSMW (N = `r sum(population$sex_behav == "MSMW")`) and transgender MSM (N = `r sum(population$gender_identity == "Transgender")`).

\
Figure \@ref(fig:PlotPop) gives a visual representation of MSM subpopulation's representation in the studies included in this paper.

\
In order to obtain the most comprehensive knowledge of the representation of MSM subpopulations in syndemic literature, we also took into account studies that do not focus on any subpopulation in particular but present disaggregated data and analyses for one or more MSM subpopulations.\
`r .N_disaggregated` studies present such data: `r .N_dis_BMSMW` for Black MSMW [@dyer2020], `r .N_dis_MSMW` for MSMW [@branstrom; @Mustanski2014; @friedman2014; @ferlatte2018], `r .N_dis_HIVMSM` for MSM living with HIV [@Kurtz2012; @ng2020] and `r .N_dis_LBMSM` for Latino and Black MSM [@Mustanski2017].

```{r plot_population, fig.cap="Number of studies focusing on a MSM subpopulation", fig.height=7, fig.id="PlotPop", fig.width=5.5, fig.align='center'}

knitr::include_graphics("./Output/Plots/Plot_population.png")
```

### Syndemic conditions and their measurement

#### Number and type of syndemic conditions identified

```{r syndemic_network_script, include=FALSE}

source("./scripts/create_syndemic_network.R", local = knitr::knit_global())

```

We identified `r nrow(nodes)` different syndemic conditions in the `r nrow(summary_quantitative)` quantitative studies we included.

The health-related conditions we identified are : depression (N = `r nodes$studies[nodes$label == "depression"]`), alcohol use disorder (AUD) (N = `r nodes$studies[nodes$label == "AUD"]`), sexual compulsivity (N = `r nodes$studies[nodes$label == "sex_comp"]`), suicidal thoughts and/or attempts (N = `r nodes$studies[nodes$label == "suicide"]`), anxiety (N = `r nodes$studies[nodes$label == "anxiety"]`), substance use disorder (SUD) (N = `r nodes$studies[nodes$label == "substance_use_disorder"]`), post-traumatic stress disorder (PTSD) (N = `r nodes$studies[nodes$label == "PTSD"]`), general mental distress (N = `r nodes$studies[nodes$label == "distress"]`), internalised homophobia (N = `r nodes$studies[nodes$label == "internalised_homophobia"]`), low self-esteem (N = `r nodes$studies[nodes$label == "esteem"]`), sexually transmitted infections (STI) (N = `r nodes$studies[nodes$label == "STI"]`), involuntary subordination (N = `r nodes$studies[nodes$label == "subordination"]`), HIV diagnosis (N = `r nodes$studies[nodes$label == "HIV"]`), hostility (N = `r nodes$studies[nodes$label == "hostility"]`), stress (N = `r nodes$studies[nodes$label == "stress"]`), experience of trauma (N = `r nodes$studies[nodes$label == "trauma"]`), alexithymia (N = `r nodes$studies[nodes$label == "alexithymia"]`), attention deficit hyperactivity disorder (ADHD) (N = `r nodes$studies[nodes$label == "ADHD"]`), cognitive escape (N = `r nodes$studies[nodes$label == "escape"]`), hypersexuality (N = `r nodes$studies[nodes$label == "hypersexuality"]`), impulsivity (N = `r nodes$studies[nodes$label == "impulsivity"]`), poor physical health (N = `r nodes$studies[nodes$label == "physical_health"]`) and sleep disturbance (N = `r nodes$studies[nodes$label == "sleep_disturbance"]`).

Alongside these health-related conditions, we also identified numerous social conditions: substance use (N = `r nodes$studies[nodes$label == "substance_use"]`), intimate partner violence (IPV) (N = `r nodes$studies[nodes$label == "IPV"]`), childhood sexual abuse (CSA) (N = `r nodes$studies[nodes$label == "CSA"]`), polysubstance use (N = `r nodes$studies[nodes$label == "polysubstance"]`), binge drinking (N = `r nodes$studies[nodes$label == "binge"]`), experiences of violence (N = `r nodes$studies[nodes$label == "violence"]`), sexual risk behaviours (N = `r nodes$studies[nodes$label == "sexual_risk"]`), alcohol use (N = `r nodes$studies[nodes$label == "alcohol_use"]`), discrimination (N = `r nodes$studies[nodes$label == "discrimination"]`), chemsex (N = `r nodes$studies[nodes$label == "chemsex"]`), loneliness (N = `r nodes$studies[nodes$label == "loneliness"]`), incarceration (N = `r nodes$studies[nodes$label == "incarceration"]`), unstable housing (N = `r nodes$studies[nodes$label == "unstable_housing"]`), tobacco use (N = `r nodes$studies[nodes$label == "tobacco_use"]`), childhood abuse (N = `r nodes$studies[nodes$label == "childhood"]`), low social support (N = `r nodes$studies[nodes$label == "social_support"]`), exchange sex (N = `r nodes$studies[nodes$label == "exchange_sex"]`), intravenous drug use (IDU) (N = `r nodes$studies[nodes$label == "IDU"]`), poverty (N = `r nodes$studies[nodes$label == "poverty"]`), unemployment (N = `r nodes$studies[nodes$label == "unemployment"]`), poor healthcare access (N = `r nodes$studies[nodes$label == "healthcare_access"]`), frequenting gay social venues (N = `r nodes$studies[nodes$label == "venues"]`) and school bullying (N = `r nodes$studies[nodes$label == "bullying"]`).

We chose to distinguish between AUD, binge drinking and alcohol use instead of merging them into a broad 'alcohol-related syndemic condition'. Some authors also considered 'heavy drinking' [@Martinez2016a; @martinez2020] or 'heavy alcohol use' [@Mimiaga2015b] but we chose to aggregate these conditions under 'binge drinking' for clarity. Similarly, we distinguish between substance use, substance use disorder, intravenous drug use, polysubstance use, marijuana use, tobacco use and chemsex since they differ in terms of potential harm and the context of use.

```{r Stall_conditions, include=FALSE}
.Stall_conditions <- summary_quantitative %>%
  filter(!is.na(CSA) | (!is.na(depression) | (!is.na(polysubstance) | (!is.na(IPV)))))

```

#### Network analysis of the syndemic conditions

In order to better understand the connections between the syndemic conditions in the studies, we elaborated a network plot (Figure \@ref(fig:SyndemicNetwork)). Nodes represent the syndemic conditions studied in the literature and edges (connectors) are drawn between two nodes when two conditions are studied in the same research paper. Edges between nodes are thicker if the two conditions are frequently studied together and the outline of the nodes represent pie chart of their frequency in syndemic literature. For example, depression, the most studied condition, appears in `r percent((nodes$studies[nodes$label == "depression"])/nrow(summary_quantitative))` of our sample. As such, the node is circled by a chart representing this proportion. Moreover, as a force-directed graph, the location of the nodes gives an indication regarding their importance in terms of frequency and linkage to other conditions. As such, depression occupies a more central position than ADHD, which was only studied once.

In this figure, we divided the conditions studied into four categories: mental health (e.g. depression, anxiety, PTSD) in blue, social conditions (e.g. substance use, loneliness) in teal, structural conditions (e.g. unemployment, healthcare access) in orange and physical health (e.g. STI, HIV) in red. The figure shows that structural syndemic conditions and physical health-related syndemic conditions are much less central, therefore much less studied, than social syndemic conditions and mental health-related syndemic conditions which are located toward the center of the figure.

Regarding the edges of the network, we identified `r nrow(edges)` pairs of syndemic conditions. However, nearly half of these pairs of conditions (N = `r nrow(.edges_1)`) appear in only one research paper. On the other hand, the most studied pairs of syndemic conditions were depression and IPV (N = `r edges$weight[edges$from == 1 & edges$to == 4]`), depression and substance use (N = (N = `r edges$weight[edges$from == 1 & edges$to == 2]`), depression and CSA (N = `r edges$weight[edges$from == 1 & edges$to == 10]`), IPV and CSA (N = `r edges$weight[edges$from == 4 & edges$to == 10]`) and depression and polysubstance use (N = `r edges$weight[edges$from == 1 & edges$to == 11]`).

```{r syndemic_network_plot, fig.cap="Network of the syndemic conditions studied in quantitative research. As a force-directed network, the most studied and interconnected conditions are the most central. Additionally, each node is circled by a pie chart of their frequency in the review. The thickness of the edges linking two nodes reflects the number of studies in which the two conditions were studied together.", fig.id="SyndemicNetwork", fig.width=7.45, fig.height=5.32, fig.align='center'}

knitr::include_graphics("./Output/Plots/syndemic_network.png")
```

#### Measurement of the syndemic conditions

```{r measurement_data, message=FALSE, warning=FALSE, include=FALSE}

source("./scripts/supplementary_tables/depression.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/IPV.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/substance.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/CSA.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/polysubstance.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/binge.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/violence.R", local = knitr::knit_global())
source("./scripts/supplementary_tables/anxiety.R", local = knitr::knit_global())

```

The next part of our research question was to examine the way the syndemic conditions were studied in the literature. For the most often studied syndemic conditions, we found a high degree of variability in their measurement. Different authors used different scales or criteria and, even when the same scale or criteria were used, the cutoff or reference period differed among studies. For the sake of brevity, we will only provide a detailed description of the measurement method for the five most studied syndemic conditions. Detailed information on the syndemic conditions most impacted by methodology heterogeneity is presented in the supplementary material, part C ( \@ref(tab:DepressionTab) to \@ref(tab:AnxietyTab)).

For depression, `r length(unique(.depression_measurement$depression_scale_criteria))` different scales or criteria are used among the `r nodes$studies[nodes$label == "depression"]` studies in which this condition was studied. The most frequently used scales are the full Center for Epidemiologic Studies-Depression scale (CESD) (N = 26) and the 10-item version of this scale (CESD-10) (N=11), the 9-item version of the Patient Health Questionnaire (PHQ-9) (N=9) and the Beck Depression Inventory (BDI-II) (N = 5).\
Cutoff for these scales varied the most for the CESD (range: 16 to 27) and the PHQ-9 (range: 5 to 15).\
When a scale is not used, the most frequent criteria to assess the presence of depression is the participant being on medication for depression in the past 12 months (N = 4).

For substance use, we looked at the class of substances explicitly screened by the authors and the reference period used. The most frequently screened class of substance is stimulants (i.e. cocaine/crack, amphetamines; N = `r .substances_prev$sum[["stimulants"]]`), followed by depressants (i.e. GHB/GBL, benzodiazepines without prescriptions; (N = `r .substances_prev$sum[["depressants"]]`), opioids (i.e. opioids misuse and heroin; N = `r .substances_prev$sum[["opioids"]]`), marijuana (N = `r .substances_prev$sum[["marijuana"]]`), ecstasy /MDMA (N = `r .substances_prev$sum[["ecstasy"]]`), hallucinogens (i.e. ketamine, psilocybin, phencyclidine; N = `r .substances_prev$sum[["hallucinogens"]]`), inhalants (nitrous oxide, popper; N = `r .substances_prev$sum[["inhalants"]]`) and new psychoactive substances (i.e. synthetic cannabinoids, cathinones; N = `r .substances_prev$sum[["NPS"]]`).\
Reference periods range from the past month to over the lifetime.

For polysubstance use, 18 studies consider that there is polysubstance use when three or more classes of substances are used while nine studies define polysubstance use as more than one class of substance being used. Furthermore, six studies exclude marijuana from the class of substances screened and two studies exclude stimulants, because stimulants were already screened in another syndemic condition.

Five types of IPV were identified: physical violence (N = `r .IPV_prevalence$sum[["physical IPV"]]`), psychological violence (N = `r .IPV_prevalence$sum[["psychological IPV"]]`), sexual violence (N = `r .IPV_prevalence$sum[["sexual IPV"]]`), gay-related violence (e.g. threats to reveal the partner's sexual orientation; N = `r .IPV_prevalence$sum[["gay-related IPV"]]`) and HIV-related violence (e.g. threats to disclose the partner's serologic status; N = `r .IPV_prevalence$sum[["HIV-related IPV"]]`). Reference periods vary from the past month to over the lifetime.

For CSA, we identified `r length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Definition"]))` different definitions and `r english(length(unique(.CSA_measurement$CSA_scale_criteria[.CSA_measurement$CSA_type == "Scale"])))` scales among the `r nodes$studies[nodes$label == "CSA"]` studies. The various definitions differ regarding the age at which sexual intercourse occurred to be qualified as childhood sexual abuse, the age gap between the victim and the perpetrator and whether the sexual intercourse was unwanted. The most frequent definition is Finkelhor's definition [@finkelhor1994]: any sexual experience before 13 years old with someone at least five years older or any sexual experience between 13 and 17 years old with someone at least ten years older (N = 7).

### Outcomes and their measurement

```{r outcomes_data, include=FALSE}

source("./scripts/create_outcomes_data.R", local = knitr::knit_global())

```

We identified `r ncol(.outcomes) - 1` different outcomes in our sample of quantitative studies. Sexual risk behaviour is the most frequently studied outcome, appearing in `r sum(!is.na(.outcomes$outcome_sexual_risk))` research papers. The second most studied outcome is HIV diagnosis (N = `r sum(!is.na(.outcomes$outcome_HIV))`), followed by STI diagnosis (N = `r sum(!is.na(.outcomes$outcome_STI))`), adherence to antiretroviral therapy (N = `r sum(!is.na(.outcomes$outcome_adherence))`), pre-exposure prophylaxis (PrEP) use (N = `r sum(!is.na(.outcomes$outcome_PrEP))`), healthcare use (N = `r sum(!is.na(.outcomes$outcome_healthcare))`) and the syndemic conditions themselves (N = `r sum(!is.na(.outcomes$outcome_syndemic))`). Less often, researchers also use viral load (N = `r sum(!is.na(.outcomes$outcome_VL))`), suicidality (N = `r sum(!is.na(.outcomes$outcome_suicide))`) or engaging in exchange sex (N = `r sum(!is.na(.outcomes$outcome_exchange_sex))`) as outcomes.

The other outcomes are only studied once: HIV screening [@Chandler2019b], HIV transmission [@satyanarayana2021], engagement in HIV care [@Biello2016], sexual violence [@semple2017], substance use [@turpin2020], engagement in chemsex [@friedman2014], having sexual intercourse with both men and women [@eaton2013] , help-seeking behaviours [@achterbergh2021], physical activity [@Zhang2019], glycaemic control [@Byg2016], hypothalamic-pituitary-adrenal (HPA) axis dysregulation [@carrico2018] and elevation in rectal cytokines/chemokines [@tapia2021].

There is a high degree of variability in the definition of what constitutes sexual risk behaviours, though the most frequently used proxy is, by far, condomless anal sex (N = `r .CAS_number`).

Used in conjunction with condomless anal sex, other proxies for sexual risk are number of partners (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_number_partners))`), substance use during sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_substance_sex))`), condomless vaginal sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_CVS))`), condomless oral sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_COS))`), sexual intercourse with female partners (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_female_sex))`), engaging in group sex (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_group_sex))`) and non-disclosure of HIV serostatus before first sexual intercourse with current main partner (N = `r sum(!is.na(.outcomes_sex_risk_measurement$outcome_non_disclosure))`).

When HIV is used as an outcome, it is self-reported in more than half of the studies (`r sum(.outcomes_measurement$measurement_outcome_HIV == "Self-report of HIV status", na.rm = T)` studies out of `r sum(!is.na(.outcomes$outcome_HIV))`) Similarly, for STI diagnosis, self-reporting is used in `r english(sum(.outcomes_measurement$measurement_outcome_STI_report == "Self-reported", na.rm = T))` papers out of `r english(sum(!is.na(.outcomes$outcome_STI)))`.

### Interventions

We identified only `r english(.N_int)` interventional studies guided by syndemic theory [@Chakrapani; @achterbergh2021]. Chakrapani et al. used a pre-test/post-test non-equivalent group design and the intervention consisted of motivational interviews. The goal of this intervention was to reduce condomless anal intercourse by enhancing condom self-efficacy and addressing co-occurring syndemic conditions. Achterbergh et al. conducted a randomised controlled trial and the intervention consisted of tailored feedback and help-seeking advice on mental health screening. The primary endpoint was to increase help-seeking behaviours; the secondary endpoints were reducing sexual risk behaviour and STI incidence. The intervention lasted for 12 months in both studies.

In Chakrapani et al. the intervention was successful in reducing the psychosocial conditions investigated (depression, alcohol use and internalised homophobia) and improving condom use. Furthermore, synergy, as measured by interaction on the additive and multiplicative scales, is present for depression and alcohol use as well as for depression and internalised homophobia on inconsistent condom use. Finally, a mediation analysis revealed that the improvement in consistent condom use was due to an improvement in condom self-efficacy caused by a reduction in alcohol use and internalised homophobia. On the other hand, the RCT conducted by Achterbergh et al. failed to impact the primary endpoint nor the two secondary endpoints.

### Summary of reviews

We identified `r english(sum(summary_reviews$design == "Systematic review"))` systematic reviews [@lewis2017; @woodward2017; @Lassiter2016] and two meta-analyses [@Rooney2018a; @Pantalone2020] in our review.

Woodward et al. aimed to identify a set of resilience resources among MSM burdened with minority stress and psychosocial condition in order to improve HIV prevention [@woodward2017]. They included `r summary_reviews$studies_included[summary_reviews$authors== "Woodward et al."]` studies in their reviews and identified 31 resilience resources, with social support and financial incomes as the two most frequently cited. Of note, most of these resources are associated with a lower HIV risk.

Lassiter et al. aimed to propose a framework to include religion and spirituality into HIV research with MSM [@Lassiter2016]. They found that religion and spirituality had a mixed effect on syndemic conditions but could be more beneficial for MSM of colour than for white MSM. Of note, the review is relatively small (including `r english(summary_reviews$studies_included[summary_reviews$authors== "Lassiter et al."])` studies) as these two factors are very rarely studied in syndemic literature.

Lewis and Wilson sought to examine the HIV prevalence and associated risk behaviours among migrants and ethnic minority MSM in North America and Europe [@lewis2017]. They found high rates of HIV prevalence and associated risk factors and propose that transnational migration could be part of a syndemic.

In their meta-analysis, Rooney et al. studied the syndemic conditions associated with sexual compulsivity among MSM and computed the mean effect size as well as whether this effect varied depending on the type of syndemic conditions associated [@Rooney2018a]. They included `r summary_reviews$studies_included[summary_reviews$authors== "Rooney et al."]` papers and found that sexual compulsivity was significantly associated with 7 syndemic conditions (anxiety, depression, CSA, alcohol use, substance use, IPV and sexual risk behaviours). The two strongest associations were with depression and anxiety.

Finally, Pantalone et al. sought to have a better understanding of the state of interventions co-targeting interconnected syndemic conditions and HIV-related health behaviours of MSM [@Pantalone2020]. They included `r summary_reviews$studies_included[summary_reviews$authors== "Pantalone et al."]` studies and found a small significant positive effect of combined behavioural interventions to improve mental health, substance use, alcohol use and sexual risk behaviours, with significant heterogeneity. Interestingly, a greater number of sessions (at least 9) and individual rather than group interventions showed greater efficacy.

### Additional frameworks used in the studies

```{r framework_data, include=FALSE}
source("./scripts/create_framework_data.R", local = knitr::knit_global())

```

Twenty studies use an additional framework alongside syndemic theory. The most widely used is resilience theory (N = 10), followed by minority stress model (N = 5), intersectionality (N=2), social-cognitive theory (N=2), salutogenesis (N=1) and ecological framework (N=1).

## How is the concept of interaction explored in syndemic research applied to MSM?

### Statistics used to show an interaction

```{r statistics_data, include=FALSE}

source("./scripts/create_statistics_data.R", local = knitr::knit_global())
```

It should be stated that, when we speak of an interaction between syndemic conditions, we do not necessarily mean 'synergy', though synergy is indeed one of the interactions we consider. Drawing from the works of authors such as Tsai and Chakrapani, we consider three types of interaction: synergistically interacting epidemics, serially causal epidemics and mutually causal epidemics [@Tsai2018a; @Chakrapani2019]. As such, statistical methods such as mediation analysis, path analysis or structural equation modelling are also considered.

\
Our review reveals a high degree of variability in the statistical analyses. The most frequently used statistical method is regression analysis using a summation score of the syndemic conditions (N = `r sum(!is.na(summary_quantitative$score))`). Among studies employing this method, `r .score_data` do not use any other method to determine the degree of interaction between syndemic conditions. In comparison, only `r .N_additivity` studies tried to evaluate departure from additivity on the additive and/or multiplicative scales, as recommended by Tsai [@tsai2015] to demonstrate synergy between syndemic conditions. Examples of measures for assessing departure from additivity are relative excess risk due to interaction (RERI), attributable proportion due to interaction (AP), and the synergy index (S). Furthermore, some studies sought to obtain a better understanding of the mechanism of interaction between the syndemic conditions and the outcome by using mediation analysis (N = `r sum(!is.na(summary_quantitative$mediation))`), moderation analysis (N = `r sum(!is.na(summary_quantitative$moderation))`), Structural Equation Modelling (N = `r sum(!is.na(summary_quantitative$SEM))`) or path analysis (N = `r sum(!is.na(summary_quantitative$path))`). Additionally, `r english(.N_latent)` studies use latent variables modelling such as Exploratory/Confirmatory Factor Analysis, Latent Profile Analysis, Latent Class Analysis or Latent Transition Analysis. Using a different approach, `r english(sum(!is.na(summary_quantitative$network)))` studies conducted by Lee and colleagues [@lee2020; @Lee] use network analysis, in which syndemic is conceptualised as a network and the syndemic conditions as interconnected nodes reinforcing each other. Finally, other statistic analyses used to assess interaction are observed/expected ratio (N = `r sum(!is.na(summary_quantitative$O_E))`), cluster analysis (N = `r sum(!is.na(summary_quantitative$cluster))`), point-biserial correlation matrix (N= `r sum(!is.na(summary_quantitative$biserial))`) and synergy factor analysis (N=1).

### Proposed mechanisms of interaction

In our sample of quantitative studies, we found only one study offering a mechanism of biological interaction [@carrico2018] and two studies offering a mechanism of bio-social interaction [@Klein2011; @tapia2021].

Carrico et al. showed that the combined effects of HIV infection and methamphetamine use are detrimental for the functioning of the Hypothalamic-Pituitary-Adrenal axis which is thought to play an important role in the reinforcing effect of stimulants.

Klein proposed that attitudes toward condom use is one of the key factors contributing to condomless sex and, subsequently, HIV infection. In their study, attitudes toward condoms are predicted by low self-esteem, as condom use represents a self-protecting mechanism, as well as by sexual preferences, substance use, race and education. Furthermore, childhood emotional neglect has a negative influence on self-esteem thus indirectly contributing to sexual risk taking.

Finally, Tapia et al. demonstrated that a higher number of syndemic conditions are associated with elevation in rectal cytokines/chemokines relevant to HIV/STI transmission. This finding presents a pathway through which depression, alcohol use disorder, substance use and PTSD may increase the biological susceptibility to HIV/STI. Unfortunately, using solely a summation score to conduct regression analysis, no synergy between the conditions is evaluated.

Qualitative studies propose mechanisms of bio-social interaction more often than quantitative studies. We identified `r english(sum(!is.na(summary_qualitative$bio_social_interaction)))` papers with such hypothesis of interaction [@Quinn; @Maiorana; @Lyons2013; @Chakrapani2019a; @Pollard2018a] . On the other hand, we did not find any qualitative paper exploring biological interaction.

Stigma and structural inequalities are indicated by the five papers as the root causes of syndemic conditions and HIV risk behaviours. Maionara et al. showed that economic dependency and fear of violence may maintain YBMSM in dysfunctional relationships plagued by IPV and substance abuse which increase their odds of incarceration and HIV infection [@Maiorana]. Furthermore, methamphetamine is used as a coping strategy while simultaneously representing a potential factor in acquiring or transmitting HIV to others, notably during sex parties.

This finding of substance use as a coping mechanism is supported by the study by Pollard et al. who found that maladaptive coping strategies to minority stress and the performative resistant space of the gay scene contributed to chemsex use [@Pollard2018a].

For Black MSM, other sources of structural violence such as racial neighbourhood segregation contribute, alongside poverty, unemployment, violence, unstable housing, incarceration and poor healthcare access to expectations surrounding masculinity which may conflict with the sexual orientation of BMSM, leading to internalised homophobia [@Quinn]. Internalised homophobia, in turn, may lead to poor mental health as well as to substance abuse and sexual risk taking as coping mechanisms. Moreover, the use of PrEP, perceived as a 'gay pill' may negatively impact the use of PrEP in this community. Lyons et al. also showed that the marginalisation of YMSM within their school, communities of origin and families resulted in a lack of gay-specific HIV prevention, education, role model and productive goal-related activities [@Lyons2013]. The HIV-positive MSM youth in this study linked these factors to their HIV acquisition. Finally, [@Chakrapani2019a] showed the role of family support and communities of MSM to serve as resilience resources to counteract the negative effects of stigma and syndemic conditions.

## What were the key findings of these studies?

In nearly all of the quantitative studies, the syndemic conditions chosen by the authors are associated with the outcomes of interest.

However, some interesting discrepancies in the findings were found. On the topic of synergy, results are conflicting between studies. The findings of Chakrapani and colleagues support the three models of interacting epidemics (synergistically interacting epidemics, serially causal epidemics and mutually causal epidemics) with the strongest support for the model of synergistically interacting epidemics [@Chakrapani2019]. These findings are supported by other studies that found synergy through departure from additivity, by using measures such as RERI, AP or S, or through a departure from multiplicativity by using significant product terms in the multiplicative scale [@Chakrapani; @chandler2020; @Chandler2019; @tomori2018; @chuang2021; @bulled2021]. However, other studies found no synergy between the syndemic conditions using the same statistical approach [@Batchelder2019; @shuper2020; @Ferlatte2018a].

Conflicting results are also present for MSMW. Brandstrom and Pachankis found no clustering of syndemic conditions for MSMW [@Branstrom2018] while Mustanski et al. found a stronger clustering of syndemic conditions in YMSMW compared to YMSM [@Mustanski2014]. Furthermore, Black MSM in the study by Dyer et al. suffered from a higher burden of STI than Black MSMW [@dyer2020]. Ferlatte et al. found more syndemic conditions in the MSM compared to the MSMW [@ferlatte2018], mainly driven by party drug use and treatment for depression or anxiety. Finally, in a sample of Latino MSMW, having two syndemic conditions predicted receptive condomless anal sex and STI diagnosis but not insertive nor vaginal condomless sex [@Munoz-Laboy2018].

In a longitudinal study [@Mustanski2017] Black MSM were found to have the highest HIV incidence while having a lower burden of syndemic condition than White MSM, raising the question of the need to focus on syndemic conditions specific to Black MSM. Interestingly, Bulled observed a synergy only for White MSM in their reanalysis of Stall's data [@bulled2021]. The authors proposed that, for racial minority MSM, structural disadvantage limiting access to healthcare may be more important than substance use and violence.

While condomless anal sex is often used as a proxy for the risk of acquiring HIV or another STI, the findings of some studies tend to show some limitation of this proxy. For example, Moeller et al. found an association between syndemic conditions and condomless anal sex but not with HIV diagnosis [@moeller]. A similar pattern appears in a longitudinal study that found an association between syndemic conditions and sexual risk behaviours but not with HIV or STI incidence [@Mustanski2017]. On the other hand, some authors found an association between syndemic conditions and HIV but not with condom use [@chuang2018].

Alcohol misuse is inconsistently linked to risk taking. Card et al. found no association between the AUDIT score and sexual risk behaviours [@Card2018] while another study using latent class analysis found that the class with alcohol misuse has a lower level of risk behaviour [@scheer2021]. Nonetheless, alcohol misuse is associated with PrEP non-adherence [@shuper2020].