### 2022-04-22 _bookdown_files/06-conclusion.Rmd

# Strengths and Limitations

This review is the first scoping review to map the current knowledge in syndemic theory applied to MSM. We charted and presented detailed data regarding the existing literature. As such, we offer insight and actionable recommendations for the future of the field. Our search strategy of electronic databases was comprehensive and developed with the help of the director of health sciences library of our institution. We combined this database search with a subsequent hand search of reference lists of the included studies. Moreover, we conducted a fully transparent and reproducible review. Other researchers may freely reuse our data and scripts to conduct their own work or use our online visualisations to identify papers pertinent to their research projects.

Nevertheless, though we took several steps to ensure the comprehensiveness of our search strategy, we may still have missed some relevant literature. Furthermore, one of our inclusion criteria was the centrality of syndemic theory to the screened paper. Although two reviewers screened a part of the studies independently, this criterion may be somewhat arbitrary and other researchers may have chosen to include papers we rejected or the contrary. Finally, the usual limitations of a scoping review apply such as the lack of risk of bias and strength of evidence assessment.

# Conclusion

Despite our critical stance on the current literature, we remain convinced that syndemic theory is a useful framework to study the health of marginalised populations in a holistic way and the current research still offers valuable insights on the health of MSM. It is vital that research continue to evolve in order to deepen our understanding and, by consequence, improve the health of marginalised populations worldwide.

# Funding

We did not receive any specific grant for this research from any funding agency in the public, commercial or not-for-profit sectors.